---
title: "EDA, Top-down, and Bottom-up Cluster Analyses"
author: "Joe Celebrezze"
date: "2023-11-24"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
here = here::here

source(here("scripts", "source_code.R"))
library(emmeans)

# Blue Mtns. Raster
blue.mtns.ras <- raster(here('data', 'blue_mtns_misc', 'bluemtnsras.tif'))
# Postfire Regen. Data (Davis et al)
postfire.regen <- raster(here('data', 'regen_prob', 'postfire_regen_prob.tif'))

wide.df <- read.csv(here('data', 'processed_data', 'wide.df.csv')) %>% 
  select(-X)
colnames(wide.df)[14:37] <- c(0:23)
long.df <- read.csv(here('data', 'processed_data', 'long.df.csv')) %>% 
  select(-X)
field.data.both <- read.csv(here('data', 'processed_data', 'field.data.both.csv')) %>% 
  select(-X)

field.data.both %>% 
  group_by(Fire) %>% 
  tally()

long.df.avg <- long.df %>% 
  group_by(relative_year) %>% 
  summarise(avg_nbr = mean(nbr), avg_delta_nbr = mean(yearly_delta_nbr))
```

# PCA (Basic)
```{r}
# Prepping Dataset
traj.metrics.df.long <- long.df %>% 
  filter(relative_year > 10) %>% 
  select(x_y, fire, postdisturbance_nbr, delta_nbr, postfire_fitted_rsq, postfire_fitted_slope, auc_delta_nbr)

traj.metrics.df.wide <- wide.df %>% 
  pivot_longer(cols = !c(x_y, Long, Lat, Plot, fire, predisturbance_nbr, postdisturbance_nbr, disturbance_nbr, delta_nbr, absolute_regrowth, relative_regrowth, max.postfire.nbr, pct_recovery, time_to_0.8recovery, time_to_0.6recovery, dnbr, hli, canopy_cover, veg_height, cwd_present, cwd_future, TE_future, TE_present, elevation, postfire_ppt, postfire_tmax, postfire_tmin, postfire_vpdmax, postfire_vpdmin), names_to = 'relative_year', values_to = 'nbr') %>% 
  mutate(relative_year = as.numeric(relative_year)) %>% 
  filter(relative_year > 10) %>% 
  select(x_y, fire, absolute_regrowth, relative_regrowth, pct_recovery)

traj.metrics.df <- merge(traj.metrics.df.long, traj.metrics.df.wide, by = c('x_y', 'fire')) %>% 
  distinct(x_y, fire, .keep_all = T) %>% 
  mutate(postfire_fitted_rsq = ifelse(is.na(postfire_fitted_rsq), 0, postfire_fitted_rsq)) %>% 
  mutate(relative_regrowth = ifelse(is.nan(relative_regrowth), NA, relative_regrowth)) %>% 
  drop_na(relative_regrowth)

traj.pca <- traj.metrics.df %>% 
  select(-x_y, -fire) %>% 
  prcomp(scale = T)

summary(traj.pca)
traj.pca$rotation[,1:2]

traj.cor <- traj.metrics.df %>% 
  select(-x_y, -fire) %>% 
  cor()

corrplot(traj.cor, method = 'number', diag = F, type = 'lower')

autoplot(traj.pca, data = traj.metrics.df,
              loadings = TRUE, loadings.colour = 'black',
              loadings.label = TRUE, loadings.label.size = 4.5) +
  theme_bw() +
  theme(axis.title = element_text(face = 'bold', size = 16))
```
traj.metrics.df, the dataframe tailored for the PCA is also going to be used in the k-means clustering analysis

# NBR Over Time (by fire)
```{r}
long.df %>% 
  ggplot(aes(x = relative_year, y = yearly_delta_nbr, color = fire)) +
    geom_point(alpha = 0.3) +
    geom_line(alpha = 0.1) +
    geom_smooth(alpha = 0.5) +
    geom_line(data = long.df.avg, aes(x = relative_year, y = avg_delta_nbr), color = 'black', linewidth = 2) +
    labs(x = 'Relative Year (Fire = 10)', y = 'NBR (Pre-fire Mean-adjusted)', color = 'Fire') +
    theme_bw()
```

# Top-Down Cluster Analysis
**Using Trajectory Metrics**
For this method, I am going to mirror the method used by Madeline when looking at the field data, but instead of inputting the field data values into the analysis, the trajectory metrics will be inputted and we'll get clusters for different 'modes' of post-fire recovery

Prepping dataset
```{r}
# Only interested in the trajectory metrics
traj.df <- traj.metrics.df %>% 
  select(-x_y, -fire, -auc_delta_nbr, -absolute_regrowth) # removing some metrics that are highly correlated

# Scaling, centering all metrics
traj.scaled.df <- as.data.frame(scale(traj.df))

# Getting rid of outliers
traj.scaled.df <- traj.scaled.df %>% 
  filter(pct_recovery < 20, pct_recovery > -10, relative_regrowth < 10)
traj.df <- traj.df %>% 
  filter(pct_recovery < 30, pct_recovery > -6, relative_regrowth < 16)
```

The below chunks are from Madeline's code
```{r}
## 3 ways to determine optimal cluster 
# 1. optimal cluster at the elbow
factoextra::fviz_nbclust(traj.scaled.df, kmeans, method = "wss") # 2 clusters

# 2. optimal cluster from gap statistic: high point!
gap_stat <- clusGap(traj.scaled.df,
                    FUN = kmeans,
                    nstart = 25,
                    K.max = 10,
                    B = 50)
# plot number of clusters vs. gap statistic
fviz_gap_stat(gap_stat) # Not sure from this one... 6?

# 3. silhouette analysis
fviz_nbclust(traj.scaled.df, kmeans, method = "silhouette") # Looks like 2
```

```{r}
#perform k-means clustering with k = 2 clusters
km <- kmeans(traj.scaled.df, centers = 2) # Note I removed nstart = 33, what does this mean?; note: I also tried k = 5

#view results
km

#plot results of final k-means model
fviz_cluster(km, data = traj.scaled.df)

#find unscaled (real-value) means of each cluster
aggregate(traj.df, by=list(cluster=km$cluster), mean)

#add cluster assigment to dataframes
traj.scaled.df <- cbind(traj.scaled.df, cluster = km$cluster)
traj.df <- cbind(traj.df, cluster = km$cluster)
wide.cluster.df <- merge(traj.df, wide.df, by = c('postdisturbance_nbr', 'delta_nbr', 'relative_regrowth', 'pct_recovery'))
```

### Time Series (Points)
```{r}
# Lengthening dataframe, adding RdNBR, scaled recovery metric, and delta NBR (deviation from prefire conditions)
long.cluster.df <- wide.cluster.df %>% 
  pivot_longer(cols = !c(x_y, Long, Lat, Plot, fire, predisturbance_nbr, postdisturbance_nbr, disturbance_nbr, delta_nbr, absolute_regrowth, relative_regrowth, postfire_fitted_rsq, postfire_fitted_slope, max.postfire.nbr, pct_recovery, time_to_0.8recovery, time_to_0.6recovery, dnbr, cluster, hli, canopy_cover, veg_height, cwd_present, cwd_future, TE_future, TE_present, elevation, postfire_ppt, postfire_tmax, postfire_tmin, postfire_vpdmax, postfire_vpdmin), names_to = 'relative_year', values_to = 'nbr') %>% 
  mutate(yearly_delta_nbr = nbr - predisturbance_nbr) %>% 
  mutate(scaled_recovery_metric = (nbr - disturbance_nbr)/(predisturbance_nbr - disturbance_nbr)) %>% 
  mutate(RdNBR = (predisturbance_nbr - nbr)/(sqrt(abs(predisturbance_nbr/1000)))) %>% 
  mutate(cluster = as.factor(cluster)) %>% 
  mutate(relative_year = as.numeric(relative_year))

# Average values
long.cluster.avg.df <- long.cluster.df %>% 
  group_by(cluster, relative_year) %>% 
    mutate(cluster = case_when(
    cluster == 1 ~ 'Fast recovery',
    cluster == 2 ~ 'Slow recovery'
  )) %>% 
  summarise(avg_nbr = mean(nbr), avg_delta_nbr = mean(yearly_delta_nbr), avg_srm = mean(scaled_recovery_metric), avg_RdNBR = mean(RdNBR)) %>% 
  drop_na(relative_year)
 

# Time series visualizations
long.cluster.df %>% 
    mutate(cluster = case_when(
    cluster == 1 ~ 'Fast recovery',
    cluster == 2 ~ 'Slow recovery'
  )) %>% 
  ggplot(aes(x = relative_year, y = nbr, color = cluster)) +
    geom_point(alpha = 0.3) +
    geom_line(alpha = 0.1) +
    geom_line(data = long.cluster.avg.df, aes(x = relative_year, y = avg_nbr, color = cluster), alpha = 0.8, linewidth = 2) +
    labs(x = 'Relative Year (Fire = 10)', y = 'NBR', color = 'Cluster') +
    theme_bw() +
    theme(axis.title = element_text(face = 'bold', size = 16),
          axis.text = element_text(size = 12),
          legend.title = element_text(face = 'bold', size = 16),
          legend.text = element_text(size = 14))
#ggsave(here('figures', 'nbr.time.series.png'), height = 9, width = 14)

long.cluster.df %>% 
    mutate(cluster = case_when(
    cluster == 1 ~ 'Fast recovery',
    cluster == 2 ~ 'Slow recovery'
  )) %>% 
    mutate(relative_year = case_when(
    cluster == 'Fast recovery' ~ relative_year + 0.075,
    cluster == 'Slow recovery' ~ relative_year - 0.075
  )) %>% 
  ggplot(aes(x = relative_year-9, y = yearly_delta_nbr, color = cluster)) +
    geom_point(alpha = 0.2) +
    geom_line(data = long.cluster.avg.df, aes(x = relative_year-9, y = avg_delta_nbr, color = cluster), alpha = 0.7, linewidth = 2.3) +
    geom_vline(xintercept = 0.5, linetype = 'dashed', linewidth = 0.4) +
    annotate(geom = 'text', x = -0.1, y = -900, label = 'Fire', size = 8, angle = 90) +
    labs(x = 'Years Post-Fire', y = 'NBR (Pre-fire Mean-adjusted)', color = 'Cluster') +
    scale_color_manual(values = c('black', 'gray70')) +
    theme_bw() +
    theme(legend.position = c(0.85, 0.12),
          axis.title = element_text(face = 'bold', size = 24),
          axis.text = element_text(size = 18),
          legend.title = element_blank(),
          axis.ticks = element_blank(),
          legend.text = element_text(size = 24))
ggsave(here('figures', 'extra_figures', 'change.in.nbr.time.series.png'), height = 9, width = 14)
```

### Time Series (Boxplots)
```{r}
significance <- c(' ', ' ', ' ', ' ', ' ', ' ', ' ',
                  ' ', ' ', ' ', ' ', ' ', ' ', ' ',
                  '*', '*', '*', '*', '*', '*', '*',
                  '*', '*', '*', '*')
relative_year <- c(0:24)
y <- c(0, 0, 0, 0, 0, 0, 0,
       0, 0, 0, 0, 0, 0, -20,
       10, 65, 120, 180, 210, 240, 260,
       280, 300, 320, 340)
y.alt <- c(0, 0, 0, 0, 0, 0, 0,
           0, 0, 0, -300, -280, -260, -250,
           -200, -150, -100, -75, -40, -10, 20,
           40, 60, 70, 80)
divergence.df <- as.data.frame(relative_year, significance, y)
                
timeseries_boxplots <- long.cluster.df %>% 
    mutate(cluster = case_when(
    cluster == 1 ~ 'Fast recovery',
    cluster == 2 ~ 'Slow recovery'
  )) %>% 
  drop_na(relative_year) %>% 
  ggplot() +
    geom_boxplot(aes(x = as.factor(relative_year), y = yearly_delta_nbr, color = cluster), alpha = 0.2, outlier.shape = NA) +
    geom_line(data = long.cluster.avg.df, aes(x = relative_year+1, y = avg_delta_nbr, color = cluster), alpha = 0.9, linewidth = 2.3) +
    #geom_segment(aes(x = 10.5, y = 410, xend = 10.5, yend = 0),
     #            arrow = arrow(length = unit(0.1, "cm")),
      #           lineend = 'square', linejoin = 'mitre',
       #          linewidth = 1.3, color = 'black') +
    geom_vline(xintercept = 10.5, linetype = 'dashed', linewidth = 0.4) +
    scale_x_discrete(labels = c('', '', '', '', '-5', '', '', '', '', '0', '', '', '', '', '5', '', '', '', '', '10', '', '', '', '')) +
    geom_text(aes(x = relative_year + 0.025, y = y.alt, label = significance),
              data = divergence.df, size = 12, fontface = 'bold') +
    annotate(geom = 'text', x = 10.1, y = -900, label = 'Fire', size = 12, angle = 90) +
    annotate(geom = 'text', x = 1, y = 300, label = 'a', size = 24, fontface = 'bold') +
    labs(x = 'Years post-fire', y = 'NBR (baseline-adjusted)', color = 'Cluster') +
    scale_color_manual(values = c('black', 'gray70')) +
    theme_bw() +
    theme(text = element_text(family = 'Arial'),
          legend.position = c(0.85, 0.12),
          axis.title = element_text(face = 'bold', size = 38),
          axis.text = element_text(size = 30),
          legend.title = element_blank(),
          axis.ticks = element_blank(),
          legend.text = element_text(size = 38),
          panel.grid.major.x = element_blank())
timeseries_boxplots
ggsave(here('figures', 'extra_figures', 'change.in.nbr.time.series.boxplots.png'), height = 9, width = 14)
```

#### Concept Figure: Defining Spectral Metrics
```{r}
concept.figure.points <- data.frame(x = c(12, 23, 10, 15, 9), y = c(437, 405, -103, 211, 452), color = c('gray50', 'gray50', 'gray30', 'gray30', 'gray30'))

long.cluster.avg.df %>% 
  filter(cluster == 'Fast recovery') %>% 
  ggplot(aes(x = relative_year-9, y = avg_nbr)) +
    geom_line(alpha = 0.7, linewidth = 2.3, color = 'black') +
    scale_x_continuous(limits = c(-9, 14.9), breaks = c(-5, 0, 5, 10)) +
    labs(x = 'Years post-fire', y = 'NBR ', color = 'Cluster') +
  # change in NBR
    geom_segment(aes(x = -9, y = 437, xend = 3, yend = 437), 
                 linetype = 'dashed', linewidth = 1, color = 'gray70') +
    geom_segment(aes(x = 1, y = 170, xend = 14.9, yend = 170), 
                 linetype = 'dashed', linewidth = 1, color = 'gray70') +
    geom_segment(aes(x = 3, y = 437, xend = 3, yend = 170),
                 linewidth = 1, color = 'gray70') +
    annotate(geom = 'text', x = 3.5, y = 300, angle = 270, size = 6,
             label = 'Change in NBR', color = 'gray70') +
    annotate(geom = 'text', x = -6, y = 457, size = 6,
             label = 'Pre-fire average NBR', color = 'gray70') + 
    annotate(geom = 'text', x = 11.9, y = 190, size = 6,
             label = 'Post-fire average NBR', color = 'gray70') + 
    # Post-fire fitted slope
    geom_segment(aes(x = 1, y = -103, xend = 13.9, yend = 405),
                 linewidth = 1, color = 'slategray') +
    annotate(geom = 'text', x = 5, y = 30, angle = 41, size = 6,
             label = 'Slope = post-fire fitted slope', color = 'slategray') +
    # Relative regrowth and % recovery
    geom_point(aes(x = x-9, y = y, color = color),
               data = concept.figure.points, size = 7) +
    scale_color_manual(values = c('gray30', 'gray50')) + 
    annotate(geom = 'text', x = 3.8, y = 437, size = 7,
             label = 'A1', color = 'gray50') + 
    annotate(geom = 'text', x = 14, y = 372, size = 7,
             label = 'A2', color = 'gray50') + 
    annotate(geom = 'text', x = -5.6, y = 190, size = 7,
             label = '% Recovery =', color = 'gray50') + 
    annotate(geom = 'text', x = -2.55, y = 190, size = 7, parse = T,
             label = 'frac(A2,A1)', color = 'gray50') + 
    annotate(geom = 'text', x = 0.1, y = -97, size = 7,
             label = 'B2', color = 'gray30') + 
    annotate(geom = 'text', x = 5.85, y = 240, size = 7,
             label = 'B3', color = 'gray30') + 
    annotate(geom = 'text', x = 0.9, y = 456, size = 7,
             label = 'B1', color = 'gray30') + 
    annotate(geom = 'text', x = -6.5, y =  100, size = 7,
             label = 'Relative regrowth =', color = 'gray30') + 
    annotate(geom = 'text', x = -1.7, y =  100, size = 7, parse = T,
             label = 'frac(B3 - B2, B1 - B2)', color = 'gray30') + 
    annotate(geom = 'text', x = -4.6, y =  25, size = 7,
             label = 'dNBR =', color = 'gray20') + 
    annotate(geom = 'text', x = -1.75, y =  25, size = 7, parse = T,
             label = 'B1 - B2', color = 'gray20') + 
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 24),
          axis.text = element_text(size = 18),
          legend.title = element_blank(),
          legend.text = element_text(size = 22),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggsave(here('figures', 'main_figures', 'conceptual.nbr.time.series.png'), height = 6, width = 11)
```

#### Blank NBR Time Series (For Presentation)
```{r}
long.cluster.avg.df %>% 
  filter(cluster == 'Fast recovery') %>% 
  ggplot(aes(x = relative_year-9, y = avg_nbr)) +
    geom_line(alpha = 1, linewidth = 2.3, color = 'black') +
    scale_x_continuous(limits = c(-9, 14), breaks = c(-5, 0, 5, 10)) +
    labs(x = 'Years post-fire', y = 'NBR ', color = 'Cluster') +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_text(face = 'bold', size = 24),
          axis.text = element_text(size = 18),
          legend.title = element_blank(),
          legend.text = element_text(size = 22),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
#ggsave(here('figures', 'blank.conceptual.nbr.time.series.png'), height = 6, width = 11)

long.cluster.avg.df %>% 
  ggplot(aes(x = relative_year-9, y = avg_nbr, color = cluster)) +
    geom_line(alpha = 1, linewidth = 2.3) +
    scale_color_manual(values = c('black', 'gray70')) +
    scale_x_continuous(limits = c(-9, 14), breaks = c(-5, 0, 5, 10)) +
    labs(x = 'Years post-fire', y = 'NBR ', color = 'Spectral recovery') +
    theme_bw() +
    theme(text = element_text(family = 'Arial'),
          legend.position = c(0.2, 0.25),
          axis.title = element_text(face = 'bold', size = 25),
          axis.text = element_text(size = 18),
          legend.title = element_text(face = 'bold', size = 25),
          legend.text = element_text(size = 22),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggsave(here('figures', 'main_figures', 'graphical_abstract', 'blank.conceptual.nbr.time.series.cluster.png'), height = 4.58, width = 11)
```

## ANOVAs/Boxplots
### When do clusters diverge?
#### T-tests
NBR
```{r}
c1 <- wide.cluster.df %>% 
  filter(cluster == 1)
c2 <- wide.cluster.df %>% 
  filter(cluster == 2)

t.test(c1$`0`, c2$`0`, var.equal = T) # sig.
t.test(c1$`1`, c2$`1`, var.equal = T) # n.s.
t.test(c1$`2`, c2$`2`, var.equal = T) # sig.
t.test(c1$`3`, c2$`3`, var.equal = T) # sig.
t.test(c1$`4`, c2$`4`, var.equal = T) # sig.
t.test(c1$`5`, c2$`5`, var.equal = T) # sig.
t.test(c1$`6`, c2$`6`, var.equal = T) # sig.
t.test(c1$`7`, c2$`7`, var.equal = T) # sig.
t.test(c1$`8`, c2$`8`, var.equal = T) # sig.
t.test(c1$`9`, c2$`9`, var.equal = T) # sig.
t.test(c1$`10`, c2$`10`, var.equal = T) # n.s.
t.test(c1$`11`, c2$`11`, var.equal = T) # n.s.
t.test(c1$`12`, c2$`12`, var.equal = T) # sig.
t.test(c1$`13`, c2$`13`, var.equal = T) # sig.
t.test(c1$`14`, c2$`14`, var.equal = T) # sig.
t.test(c1$`15`, c2$`15`, var.equal = T) # sig.
t.test(c1$`16`, c2$`16`, var.equal = T) # sig.
t.test(c1$`17`, c2$`17`, var.equal = T) # sig.
t.test(c1$`18`, c2$`18`, var.equal = T) # sig.
t.test(c1$`19`, c2$`19`, var.equal = T) # sig.
t.test(c1$`20`, c2$`20`, var.equal = T) # sig.
t.test(c1$`21`, c2$`21`, var.equal = T) # sig.
t.test(c1$`22`, c2$`22`, var.equal = T) # sig.
t.test(c1$`23`, c2$`23`, var.equal = T) # sig.

rm(c1, c2)
```

Yearly Delta NBR
```{r}
for(i in 0:23){
c1 <- long.cluster.df %>% 
  filter(cluster == 1 & relative_year == i)
c2 <- long.cluster.df %>% 
  filter(cluster == 2 & relative_year == i)
print(i)
print(t.test(c1$yearly_delta_nbr, c2$yearly_delta_nbr, var.equal = T))
}
```

#### RANOVA
```{r}
# Option 1
diverge.df <- long.cluster.df %>% 
  filter(relative_year > 9)

diverge.mod <- lmer(nbr ~ relative_year*cluster + (1|fire), data = diverge.df)
summary(diverge.mod)

# Option 2
for(i in 9:23) {
  df <- long.cluster.df %>% 
    filter(relative_year == i)
  mod <- lmer(yearly_delta_nbr ~ cluster + (1|fire), df)
  print(i)
  print(summary(mod))
}
```

### Differences in Fire
```{r}
# Differences in fire
fire.lm <- lm(cluster ~ fire, data = wide.cluster.df)
anova(fire.lm)
fire.aov <- aov(fire.lm)
TukeyHSD(fire.aov)
anova_labels <- data.frame(fire=c(1:8), label=c('a', 'a', 'a', 'ab', 'bc', 'bc', 'bc', 'c'))

wide.cluster.df %>% 
 mutate(fire=factor(fire)) %>% 
 mutate(fire = case_when(
   fire == "Easy" ~ "Easy",
   fire == "HR" ~ "Hash \nRock",
   fire == "School" ~ "School",
   fire == "Wheeler" ~ "Wheeler \nPoint",
   fire == "Flagtail" ~ "Flagtail",
   fire == "Calamity" ~ "Calamity \nComplex",
   fire == "BC" ~ "Bridge \nCreek",
   fire == "Egley" ~ "Egley \nComplex"
 )) %>% 
 mutate(cluster = case_when(
   cluster == 1 ~ "Fast \nRecovery",
   cluster == 2 ~ "Slow \nRecovery")) %>%  
 mutate(fire=fct_relevel(fire,c("Easy", "Hash \nRock", "School", "Wheeler \nPoint", "Flagtail", 
                                "Calamity \nComplex", "Bridge \nCreek", "Egley \nComplex"))) %>%
 ggplot(aes(x = fire, y = cluster)) +
  geom_jitter(height = 0.3, alpha = 0.6) +
  #scale_y_continuous(breaks = c(1, 2)) +
  labs(x = 'Fire', y = 'Spectral Trajectory Cluster') +
  geom_text(aes(x = fire, y = 1.5, label = label), data = anova_labels, fontface = 'italic', size = 5) +
  theme_bw() +
  theme(axis.title = element_blank(),
        #axis.title = element_text(face = 'bold', size = 16),
        axis.text.x = element_text(size = 15, face = 'bold'),
        axis.text.y = element_text(size = 15, face = 'bold', angle = 90, hjust = 0.5),
        axis.ticks = element_blank())
ggsave(here('figures', 'extra_figures', 'fire.vs.cluster.png'), height = 5, width = 9.5)

diff.by.fire <- wide.cluster.df %>% 
  group_by(fire, cluster) %>% 
  tally() %>% 
  pivot_wider(names_from = cluster, values_from = n) %>% 
  mutate(cluster_ratio = `1`/`2`)
diff.by.fire %>% 
  ungroup() %>% 
  summarise(one = sum(`1`), two = sum(`2`))
```

### Climate and Topography Normals
#### Data Wrangling
Because the PRSIM 30yr normals data was too large for me to push into GitHub from my local server, I had to pull in a csv file with the climate normals already combined with the clustered dataset (at the bottom of this chunk); I commented out the code to show how, if the data was available, the code ran prior to this workaround
```{r}
# PRISM 30yr normals, 800m resolution
#precip.ras <- raster(here('data', 'climate', 'precip_30yr_normal', 'PRISM_ppt_30yr.asc'))
#tmean.ras <- raster(here('data', 'climate', 'tmean_30yr_normal', 'PRISM_tmean_30yr.asc'))
#vpdmax.ras <- raster(here('data', 'climate', 'vpdmax_30yr_normal', 'PRISM_vpdmax_30yr.asc'))
# Current CRS: +proj=longlat +datum=NAD83 +no_defs; needs to be changed to +proj=longlat +datum=WGS84 +no_defs
#precip.ras <- projectRaster(from = precip.ras, to = blue.mtns.ras, crs = #crs(postfire.regen))
#tmean.ras <- projectRaster(from = tmean.ras, to = blue.mtns.ras, crs = #crs(postfire.regen))
#vpdmax.ras <- projectRaster(from = vpdmax.ras, to = blue.mtns.ras, crs = crs(postfire.regen))

# Extract Climate Variables from Rasters to add to wide.cluster.df
wide_cluster_coords <- wide.cluster.df[,c(9,10)] # define coordinates
wide_cluster_pts <- wide.cluster.df %>% 
  select(Long, Lat) %>% 
  SpatialPointsDataFrame(coords = wide_cluster_coords, proj4string = crs(postfire.regen)) # transform to spatial points dataframe
## Precip ##
#precip_vals <- as.data.frame(extract(precip.ras, wide_cluster_pts))
#colnames(precip_vals)[1] <- 'precip' # renaming column
## Mean Temp ##
#tmean_vals <- as.data.frame(extract(tmean.ras, wide_cluster_pts))
#colnames(tmean_vals)[1] <- 'tmean' # renaming column
## Max VPD ##
#vpdmax_vals <- as.data.frame(extract(vpdmax.ras, wide_cluster_pts))
#colnames(vpdmax_vals)[1] <- 'vpdmax' # renaming column

## Postfire Regen Probability ##
#regen_vals <- as.data.frame(extract(postfire.regen, wide_cluster_pts))
#colnames(regen_vals)[1] <- 're-0gen_prob' # renaming column

# Combining with wide.cluster.df
#climate_vals <- cbind(wide_cluster_coords, precip_vals, tmean_vals, vpdmax_vals, regen_vals)
#wide.cluster.df <- merge(wide.cluster.df, climate_vals, by = c('Long', 'Lat')) %>% 
    #distinct(x_y, .keep_all = T) # removing duplicates
wide.cluster.df <- read.csv(here('data', 'processed_data', 'clusters.and.climate.normals.csv'))
```

#### Stats Tests
Precip
```{r}
# Normal distribution?
precip.lm <- lm(data = wide.cluster.df, precip ~ cluster)
plot(precip.lm) # CLT applies, but qqplot looks kind funky
# Equal variances?
var.test(data = wide.cluster.df, precip ~ cluster) # No, variances not equal
# T-test
c1 <- wide.cluster.df %>% 
  filter(cluster == 1) %>% 
  select(precip)
c2 <- wide.cluster.df %>% 
  filter(cluster == 2) %>% 
  select(precip)
t.test(c1, c2, var.equal = F) # significant differences
```

Mean Temp.
```{r}
# Normal distribution?
tmean.lm <- lm(data = wide.cluster.df, tmean ~ cluster)
plot(tmean.lm) # CLT applies, but qqplot looks kind funky (basically looks the same as precip)
# Equal variances?
var.test(data = wide.cluster.df, tmean ~ cluster) # No, variances not equal
# T-test
c1 <- wide.cluster.df %>% 
  filter(cluster == 1) %>% 
  select(tmean)
c2 <- wide.cluster.df %>% 
  filter(cluster == 2) %>% 
  select(tmean)
t.test(c1, c2, var.equal = F) # significant differences
```

Max VPD
```{r}
# Normal distribution?
vpdmax.lm <- lm(data = wide.cluster.df, vpdmax ~ cluster)
plot(vpdmax.lm)
# Equal variances?
var.test(data = wide.cluster.df, vpdmax ~ cluster) # No, variances not equal
# T-test
c1 <- wide.cluster.df %>% 
  filter(cluster == 1) %>% 
  select(vpdmax)
c2 <- wide.cluster.df %>% 
  filter(cluster == 2) %>% 
  select(vpdmax)
t.test(c1, c2, var.equal = F) # significant differences
```

CHILI
```{r}
# Normal distribution?
chili.lm <- lm(data = wide.cluster.df, hli ~ cluster)
plot(chili.lm) # more of a uniform distribution; CLT still applies
# Equal variances?
var.test(data = wide.cluster.df, hli ~ cluster) # No, variances not equal
# T-test
c1 <- wide.cluster.df %>% 
  filter(cluster == 1) %>% 
  select(hli)
c2 <- wide.cluster.df %>% 
  filter(cluster == 2) %>% 
  select(hli)
t.test(c1, c2, var.equal = F) # significant differences
```

#### Boxplots
```{r}
wide.cluster.df %>% 
  mutate(cluster = as.factor(cluster)) %>% 
  pivot_longer(cols = c(precip, tmean, vpdmax, hli), names_to = 'climate_var', values_to = 'climate_val') %>% 
  ggplot(aes(x = cluster, y = climate_val, color = cluster)) +
    geom_boxplot(outlier.shape = NA, size = 0.8) +
    geom_jitter(width = 0.25, alpha = 0.05) +
    labs(x = 'Spectral Trajectory Cluster') +
    facet_wrap(~climate_var, scales = 'free', labeller = as_labeller(climate_var_names), ncol = 4) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title.y = element_blank(),
          axis.title.x = element_text(face = 'bold', size = 16),
          axis.text.x = element_text(size = 14),
          axis.text.y = element_text(size = 12),
          strip.text = element_text(face = 'bold', size = 12))
ggsave(here('figures', 'extra_figures', 'climate.vs.cluster.boxplots.png'), height = 7, width = 15)
```


### Clusters vs. Field Data
Getting Combined Dataset Ready
```{r}
field.data.both <- field.data.both %>% 
  drop_na(Long)

main.df <- merge(wide.cluster.df, field.data.both, by = 'Plot') %>% 
  select(-Long.y, -Lat.y, -Fire) %>% 
  rename(Long = Long.x, Lat = Lat.x) %>% 
  mutate(aspect = as.numeric(aspect),
         cluster = as.factor(cluster),
         log.seedling.density = log(seedling.density),
         log.regen.density = log(regen.density)) %>% 
  mutate(log.seedling.density = ifelse(log.seedling.density < 0, 0, log.seedling.density),
         log.regen.density = ifelse(log.regen.density < 0, 0, log.regen.density)) %>%  # making any -Inf values for log.seedling.density and log.regen.density = 0
  distinct() # omitting duplicate values

main.df %>% 
  group_by(cluster) %>% 
  tally()
```

#### Stats Tests

##### RANOVA
```{r}
summary(lmer(dnbr ~ cluster + (1|fire), main.df)) # p = 0.0364
summary(lmer(postfire_fitted_slope ~ cluster + (1|fire), main.df)) # sig.
summary(lmer(delta_nbr ~ cluster + (1|fire), main.df)) # sig.
summary(lmer(pct_recovery ~ cluster + (1|fire), main.df)) # sig.
summary(lmer(relative_regrowth ~ cluster + (1|fire), main.df)) # sig.
summary(lmer(densiometer ~ cluster + (1|fire), main.df)) # n.s.
summary(lmer(regen.density ~ cluster + (1|fire), main.df)) # n.s.
summary(lmer(seedling.pct ~ cluster + (1|fire), main.df)) # n.s.
summary(lmer(shrub.pct ~ cluster + (1|fire), main.df)) # sig.
summary(lmer(grass.pct ~ cluster + (1|fire), main.df)) # n.s.
summary(lmer(herb.pct ~ cluster + (1|fire), main.df)) # n.s.
summary(lmer(slope ~ cluster + (1|fire), main.df)) # n.s.
summary(lmer(elevation ~ cluster + (1|fire), main.df)) # n.s.
```

##### T-test
Looking at two clusters, t-test

% Grass
```{r}
# Normal distribution?
grass.lm <- lm(data = main.df, grass.pct ~ cluster)
plot(grass.lm)
# Equal variances?
var.test(data = main.df, grass.pct ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(grass.pct)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(grass.pct)
t.test(c1, c2, var.equal = T)
```

% herb
```{r}
# Normal distribution?
herb.lm <- lm(data = main.df, herb.pct ~ cluster)
plot(herb.lm)
# Equal variances?
var.test(data = main.df, herb.pct ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(herb.pct)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(herb.pct)
t.test(c1, c2, var.equal = T)
```
**Significant difference**

% shrub
```{r}
# Normal distribution?
shrub.lm <- lm(data = main.df, shrub.pct ~ cluster)
plot(shrub.lm)
# Equal variances?
var.test(data = main.df, shrub.pct ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(shrub.pct)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(shrub.pct)
t.test(c1, c2, var.equal = T)
```
**Significant difference**

% seedling
```{r}
# Normal distribution?
seedling.lm <- lm(data = main.df, seedling.pct ~ cluster)
plot(seedling.lm) # NOT A NORMAL DISTRIBUTION, but hopefully central limit theorem applies (n > 30 for each group)
# Equal variances?
var.test(data = main.df, seedling.pct ~ cluster) # NO! Indicate in t.test below
# T-test (Welch's)
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(seedling.pct)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(seedling.pct)
t.test(c1, c2, var.equal = F)
```

% green
```{r}
# Normal distribution?
green.lm <- lm(data = main.df, green.pct ~ cluster)
plot(green.lm)
# Equal variances?
var.test(data = main.df, green.pct ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(green.pct)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(green.pct)
t.test(c1, c2, var.equal = T)
```
**significant differences**

% npv
```{r}
# Normal distribution?
npv.lm <- lm(data = main.df, npv.pct ~ cluster)
plot(npv.lm)
# Equal variances?
var.test(data = main.df, npv.pct ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(npv.pct)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(npv.pct)
t.test(c1, c2, var.equal = T)
```

Log Seedling Density
```{r}
# Normal distribution?
log.seedling.lm <- lm(data = main.df, log.seedling.density ~ cluster)
plot(log.seedling.lm)
# Equal variances?
var.test(data = main.df, log.seedling.density ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(log.seedling.density)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(log.seedling.density)
t.test(c1, c2, var.equal = T)
```

Log Regen Density
```{r}
# Normal distribution?
log.regen.lm <- lm(data = main.df, log.regen.density ~ cluster)
plot(log.regen.lm)
# Equal variances?
var.test(data = main.df, log.regen.density ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(log.regen.density)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(log.regen.density)
t.test(c1, c2, var.equal = T)
```

Slope
```{r}
# Normal distribution?
slope.lm <- lm(data = main.df, slope ~ cluster)
plot(slope.lm)
# Equal variances?
var.test(data = main.df, slope ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(slope)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(slope)
t.test(c1, c2, var.equal = T)
```

Aspect
```{r}
# Normal distribution?
aspect.lm <- lm(data = main.df, aspect ~ cluster)
plot(aspect.lm) # Doesn't look great, but hopefully CLT applies
# Equal variances?
var.test(data = main.df, aspect ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(aspect)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(aspect)
t.test(c1, c2, var.equal = T)
```
**significant differences**

Densiometer
```{r}
# Normal distribution?
densio.lm <- lm(data = main.df, densiometer ~ cluster)
plot(densio.lm)
# Equal variances?
var.test(data = main.df, densiometer ~ cluster)
# T-test
c1 <- main.df %>% 
  filter(cluster == 1) %>% 
  select(densiometer)
c2 <- main.df %>% 
  filter(cluster == 2) %>% 
  select(densiometer)
t.test(c1, c2, var.equal = T)
```
**significant differences**

#### Boxplots
##### Wrangling
```{r}
# T-test Results
cluster <- c(rep(c('Fast \nrecovery', 'Slow \nrecovery'), 8))
cluster.no <- c(rep(c(0.6, 2), 8))
var <- c(rep('delta_nbr', 2), rep('pct_recovery', 2),
               rep('postfire_fitted_slope', 2), rep('relative_regrowth', 2),
               rep('regen.density', 2), rep('seedling.pct', 2),
               rep('shrub.pct', 2), rep('grass.pct', 2))
y.val <- c(rep(7, 2), rep(1.73, 2),
           rep(50, 2), rep(3.7, 2), 
           rep(9, 2), rep(64, 2),
           rep(85, 2), rep(70, 2))
panel.y.val <- c(rep(9, 2), rep(1.8, 2),
           rep(60, 2), rep(4.4, 2), 
           rep(18500, 2), rep(63, 2),
           rep(88, 2), rep(85, 2))
label <- c('*', '', '*', '', '*', '', '*', '', '', '', '', '', '*', '', '', '*')
panel.label <- c('b', '', 'c', '', 'd', '', 'e', '', 'f', '', 'g', '', 'h', '', 'i', '')
t.long.cluster.df <- data.frame(cluster, cluster.no, var, y.val, label, panel.y.val, panel.label) 

df <- main.df %>% 
  mutate(cluster = case_when(
    cluster == 1 ~ 'Fast \nrecovery',
    cluster == 2 ~ 'Slow \nrecovery'
  )) %>% 
  pivot_longer(cols = c(pct_recovery, postfire_fitted_slope, relative_regrowth, delta_nbr, regen.density, seedling.pct, shrub.pct, grass.pct), values_to = 'val', names_to = 'var')
```

##### Linear Y-Axis
```{r}
vars <- c('delta_nbr', 'pct_recovery', 'postfire_fitted_slope',
                       'relative_regrowth', 'seedling.pct', 
                       'shrub.pct', 'grass.pct')

plot_list <- list()

for(i in 1:length(vars)){
text.df <- t.long.cluster.df %>% 
  filter(var == vars[i])

plot <- df %>% 
  filter(var == vars[i]) %>% 
  ggplot(aes(x = cluster, y = val, color = cluster)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.8) +
    geom_jitter(width = 0.1, height = 0, alpha = 0.1) +
    geom_text(aes(x = cluster, y = y.val, label = label),
              color = 'black', size = 14, fontface = 'bold',
              data = text.df)  +
    geom_text(aes(x = cluster.no, y = panel.y.val, label = panel.label),
              color = 'black', size = 18, fontface = 'bold',
              data = text.df)  +
    scale_color_manual(values = c('black', 'gray70')) +
    labs(x = 'Spectral recovery cluster', y = ' ') +
    facet_wrap(~var, scales = 'free',
               labeller = as_labeller(spectral_field_var_names),
               strip.position = 'left') +
    theme_bw() +
    theme(text = element_text(family = 'Arial'),
          legend.position = 'none',
          axis.title = element_blank(),
          axis.text.x = element_text(face = 'bold', size = 22),
          axis.text.y = element_text(size = 18),
          strip.text = element_text(face = 'bold', size = 24),
          strip.background = element_blank(),
          strip.placement = 'outside',
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
plot_list[[i]] <- plot
}
```

##### Log-Transformed Y-Axis
```{r}
text.df <- t.long.cluster.df %>% 
  filter(var == 'regen.density')

plot <- df %>% 
  filter(var == 'regen.density') %>% 
  ggplot(aes(x = cluster, y = val, color = cluster)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.8) +
    geom_jitter(width = 0.1, height = 0, alpha = 0.1) +
    geom_text(aes(x = cluster, y = y.val, label = label),
              color = 'black', size = 14, fontface = 'bold',
              data = text.df)  +
    geom_text(aes(x = cluster.no, y = panel.y.val, label = panel.label),
              color = 'black', size = 18, fontface = 'bold',
              data = text.df)  +
    scale_color_manual(values = c('black', 'gray70')) +
    labs(x = 'Spectral recovery cluster', y = ' ') +
    facet_wrap(~var, scales = 'free',
               labeller = as_labeller(spectral_field_var_names),
               strip.position = 'left') +
    scale_y_continuous(trans = 'log', breaks = c(50, 1000, 20000)) +
    theme_bw() +
    theme(text = element_text(family = 'Arial'),
          legend.position = 'none',
          axis.title = element_blank(),
          axis.text.x = element_text(face = 'bold', size = 22),
          axis.text.y = element_text(size = 18),
          strip.text = element_text(face = 'bold', size = 24, hjust = 1),
          strip.background = element_blank(),
          strip.placement = 'outside',
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
plot_list[[8]] <- plot

#rm(df, plot, text.df)
```

##### Arranged
```{r, warning=FALSE}
spectral_boxplots <- cowplot::plot_grid(plot_list[[1]], plot_list[[2]], plot_list[[3]],
                                        plot_list[[4]], plot_list[[8]], plot_list[[5]],
                                        plot_list[[6]], plot_list[[7]], ncol = 4)
spectral_boxplots

rm(plot_list)
```

Looking at spectral trajectory cluster vs. field data
```{r}
main.df %>% 
  pivot_longer(cols = c(grass.pct, herb.pct, shrub.pct, seedling.pct, green.pct, npv.pct, log.seedling.density, log.regen.density, densiometer), values_to = 'field_val', names_to = 'field_var')  %>% 
  ggplot(aes(x = cluster, y = field_val, color = cluster)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.3) +
    labs(x = 'Spectral Trajectory Cluster') +
    facet_wrap(~field_var, scales = 'free', labeller = as_labeller(field_var_names)) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title.y = element_blank(),
          axis.title.x = element_text(face = 'bold', size = 16),
          axis.text.x = element_text(size = 14),
          axis.text.y = element_text(size = 12),
          strip.text = element_text(face = 'bold', size = 12))
ggsave(here('figures', 'extra_figures', 'field.vs.clusters.boxplots.png'), height = 8, width = 9)
```

Canopy cover, % shrub, % green, % herb and aspect show a significant difference between clusters; however, I think the patterns we observed in the field could be somewhat hidden, as there were distinct differences observed in the 'trailing edge' vs. the 'forested' sites

## Top-Down Figure (Arranged)
```{r, warning = FALSE}
topdown1 <- cowplot::plot_grid(timeseries_boxplots, spectral_boxplots, rel_heights = c(11, 9), ncol = 1)
topdown1
ggsave(here('figures', 'main_figures', 'top.down.figure.png'), width = 21, height = 20)
```

# -------------------
# Bottom-Up Cluster Analysis

## K-Means Cluster Analysis
**Using Field Data**

Prepping dataset
```{r}
# Only interested in the field data metrics
field.df <- main.df %>% 
  select(grass.pct, herb.pct, shrub.pct, seedling.pct, green.pct, npv.pct, log.regen.density, densiometer) %>%  # trying log regen density rather than log seedling density
  drop_na(seedling.pct)

field.df <- main.df %>% 
  select(grass.pct, shrub.pct, seedling.pct, log.regen.density) %>%  # trying log regen density rather than log seedling density
  drop_na(log.regen.density)

# Scaling, centering all metrics
field.scaled.df <- as.data.frame(scale(field.df))
```

The below chunks are from Madeline's code
```{r}
## 3 ways to determine optimal cluster 
# 1. optimal cluster at the elbow
fviz_nbclust(field.scaled.df, kmeans, method = "wss") # 2, 4 or 7 clusters

# 2. optimal cluster from gap statistic: high point!
gap_stat <- clusGap(field.scaled.df,
                    FUN = kmeans,
                    nstart = 25,
                    K.max = 10,
                    B = 50)
# plot number of clusters vs. gap statistic
fviz_gap_stat(gap_stat) # Unclear

# 3. silhouette analysis
fviz_nbclust(field.scaled.df, kmeans, method = "silhouette") # Looks like 3
```

```{r}
#perform k-means clustering with k = 3 clusters
km <- kmeans(field.scaled.df, centers = 3) # Note I removed nstart = 33, what does this mean?; note: I also tried k = 5

#view results
km

#plot results of final k-means model
fviz_cluster(km, data = field.scaled.df)

#find unscaled (real-value) means of each cluster
aggregate(field.df, by=list(cluster=km$cluster), mean)

#add cluster assigment to dataframes
field.scaled.df <- cbind(field.scaled.df, cluster = km$cluster)
field.df <- cbind(field.df, field.cluster = km$cluster)
field.cluster.df <- merge(field.df, main.df, by = c('grass.pct', 'shrub.pct', 'log.regen.density', 'seedling.pct')) %>% 
  distinct(Plot, .keep_all = T) #%>% # removing duplicates
  #rename(green.pct = green.pct.x, npv.pct = npv.pct.x) %>% 
  #select(-green.pct.y, -npv.pct.y)
```

## Visualizations
```{r}
# Lengthening dataframe, adding RdNBR, scaled recovery metric, and delta NBR (deviation from prefire conditions)
colnames(field.cluster.df)[27:50] <- c(0:23)

long.field.cluster.df <- field.cluster.df %>% 
  pivot_longer(cols = c(`0`, `1`, `2`, `3`, `4`, `5`, `6`, `7`, `8`, `9`, `10`,
                        `11`, `12`, `13`, `14`, `15`, `16`, `17`, `18`, `19`, `20`,
                        `21`, `22`, `23`), names_to = 'relative_year', values_to = 'nbr') %>% 
  mutate(yearly_delta_nbr = nbr - predisturbance_nbr) %>% 
  mutate(scaled_recovery_metric = (nbr - disturbance_nbr)/(predisturbance_nbr - disturbance_nbr)) %>% 
  mutate(RdNBR = (predisturbance_nbr - nbr)/(sqrt(abs(predisturbance_nbr/1000)))) %>% 
  mutate(field.cluster = as.factor(field.cluster)) %>% 
  mutate(relative_year = as.numeric(relative_year))
# Average values
long.field.cluster.avg.df <- long.field.cluster.df %>%
  mutate(field.cluster = case_when(field.cluster == 1 ~ 'Shrub-dominated',
                                   field.cluster == 2 ~ 'Tree-dominated',
                                   field.cluster == 3 ~ 'Grass-dominated')) %>% 
  group_by(field.cluster, relative_year) %>% 
  summarise(avg_nbr = mean(nbr), avg_delta_nbr = mean(yearly_delta_nbr), avg_srm = mean(scaled_recovery_metric), avg_RdNBR = mean(RdNBR))

# Number of sites per cluster
field.cluster.df %>% 
  group_by(field.cluster) %>% 
  tally()

# Time series visualizations
long.field.cluster.df %>% 
  mutate(field.cluster = case_when(field.cluster == 1 ~ 'Shrub-dominated',
                                   field.cluster == 2 ~ 'Tree-dominated',
                                   field.cluster == 3 ~ 'Grass-dominated')) %>% 
  ggplot(aes(x = relative_year, y = nbr, color = field.cluster)) +
    geom_point(alpha = 0.3) +
    geom_line(alpha = 0.1) +
    geom_line(data = long.field.cluster.avg.df, aes(x = relative_year, y = avg_nbr, color = field.cluster), alpha = 0.8, linewidth = 2) +
    labs(x = 'Relative year (fire = 10)', y = 'NBR', color = 'Field cluster') +
    theme_bw() +
    theme(axis.title = element_text(face = 'bold', size = 16),
          axis.text = element_text(size = 12),
          legend.title = element_text(face = 'bold', size = 16),
          legend.text = element_text(size = 14))
#ggsave(here('figures', 'field.cluster.nbr.time.series.png'), height = 9, width = 14)

long.field.cluster.df %>% 
  mutate(field.cluster = case_when(field.cluster == 1 ~ 'Shrub-dominated',
                                   field.cluster == 2 ~ 'Tree-dominated',
                                   field.cluster == 3 ~ 'Grass-dominated')) %>% 
  ggplot(aes(x = relative_year-10, y = yearly_delta_nbr, color = field.cluster)) +
    geom_point(alpha = 0.3) +
    geom_line(alpha = 0.1) +
    geom_line(data = long.field.cluster.avg.df, aes(x = relative_year-10, y = avg_delta_nbr, color = field.cluster), alpha = 0.8, linewidth = 2) +
    labs(x = 'Relative year (fire = 10)', y = 'NBR (pre-fire mean-adjusted)', color = 'Field cluster') +
    theme_bw() +
    theme(axis.title = element_text(face = 'bold', size = 16),
          axis.text = element_text(size = 12),
          legend.text = element_text(face = 'bold', size = 16),
          legend.title = element_blank())
#ggsave(here('figures', 'field.cluster.change.in.nbr.time.series.png'), height = 9, width = 14)
```

### Main Time Series
```{r}
significance <- c(' ', ' ', ' ', ' ', ' ', ' ', ' ',
                  ' ', ' ', ' ', ' ', ' ', ' ', ' ',
                  ' ', 's', 's', 's', 's', 's', 's',
                  's', 's', 'st', 'st')
relative_year <- c(0:24)
y <- c(0, 0, 0, 0, 0, 0, 0,
       0, 0, 0, 0, 0, 0, -20,
       10, 65, 120, 180, 210, 240, 260,
       280, 300, 290, 280)
y.alt <- c(0, 0, 0, 0, 0, 0, 0,
           0, 0, 0, -300, -280, -260, -250,
           -200, -150, -100, -75, -40, -10, 20,
           40, 60, 70, 80)
divergence.df <- as.data.frame(relative_year, significance, y)

field.cluster.df %>% 
  group_by(field.cluster) %>% 
  tally()

field.cluster.time.series <- long.field.cluster.df %>% 
  mutate(field.cluster = case_when(field.cluster == 1 ~ 'Shrub-dominated',
                                   field.cluster == 2 ~ 'Tree-dominated',
                                   field.cluster == 3 ~ 'Grass-dominated')) %>% 
  drop_na(relative_year) %>% 
  ggplot() +
    geom_boxplot(aes(x = as.factor(relative_year), y = yearly_delta_nbr, color = field.cluster), alpha = 0.2, outlier.shape = NA) +
    geom_line(data = long.field.cluster.avg.df, aes(x = relative_year+1, y = avg_delta_nbr, color = field.cluster), alpha = 0.9, linewidth = 2.3) +
    geom_vline(xintercept = 10.5, linetype = 'dashed', linewidth = 0.4) +
    geom_text(aes(x = relative_year + 0.025, y = y, label = significance),
              data = divergence.df, size = 8.5, fontface = 'bold') +
    scale_x_discrete(labels = c('', '', '', '', '-5', '', '', '', '', '0', '', '', '', '', '5', '', '', '', '', '10', '', '', '', '')) +
    scale_y_continuous(breaks = c(-900, -600, -300, 0, 300)) +
    annotate(geom = 'text', x = 10.1, y = -900, label = 'Fire', size = 12, angle = 90) +
    annotate(geom = 'text', x = 1, y = 300, label = 'a', size = 24, fontface = 'bold') +
    labs(x = 'Years post-fire', y = 'NBR (baseline-adjusted)', color = 'Cluster') +
    scale_color_manual(values = c('#DDC49B', '#002F00', '#08A045')) +
    theme_bw() +
    theme(text = element_text(family = 'Arial'),
          legend.position = c(0.85, 0.12),
          axis.title.x = element_text(face = 'bold', size = 38),
          axis.title.y = element_text(margin = margin(r = 20), face = 'bold', size = 38),
          axis.text = element_text(size = 30),
          legend.title = element_blank(),
          axis.ticks = element_blank(),
          legend.text = element_text(size = 38),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank())
field.cluster.time.series
ggsave(here('figures', 'extra_figures', 'field.cluster.nbr.time.series.boxplots.png'), height = 9, width = 14)
```

### Boxplots
#### Exploratory
Looking at field data cluster vs. field data
```{r}
field.cluster.df %>% 
  mutate(field.cluster = case_when(field.cluster == 1 ~ 'Shrub\ndominated',
                                   field.cluster == 2 ~ 'Tree\ndominated',
                                   field.cluster == 3 ~ 'Grass\ndominated')) %>% 
  pivot_longer(cols = c(grass.pct, herb.pct, shrub.pct, seedling.pct, green.pct, npv.pct, log.regen.density, densiometer, log.seedling.density), values_to = 'field_val', names_to = 'field_var')  %>% 
  ggplot(aes(x = field.cluster, y = field_val, color = field.cluster)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.3) +
    scale_color_manual(values = c('#DDC49B', '#002F00', '#08A045')) +
    labs(x = 'Field data cluster', y = ' ') +
    facet_wrap(~field_var, scales = 'free', labeller = as_labeller(field_var_names)) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_blank(),
          axis.text.x = element_text(face = 'bold', size = 14),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(face = 'bold', size = 14))
ggsave(here('figures', 'extra_figures', 'field.cluster.field.boxplots.png'), height = 10, width = 12)
```

vs. Spectral Metrics
```{r}
field.cluster.df %>% 
  mutate(field.cluster = case_when(field.cluster == 1 ~ 'Shrub\ndominated',
                                   field.cluster == 2 ~ 'Tree\ndominated',
                                   field.cluster == 3 ~ 'Grass\ndominated')) %>% 
  pivot_longer(cols = c(postfire_fitted_slope, relative_regrowth, delta_nbr, pct_recovery), values_to = 'spectral_val', names_to = 'spectral_var')  %>% 
  ggplot(aes(x = field.cluster, y = spectral_val, color = field.cluster)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.3) +
    scale_color_manual(values = c('tan', 'darkgreen', 'forestgreen')) +
    labs(x = 'Field data cluster', y = ' ') +
    facet_wrap(~spectral_var, scales = 'free', labeller = as_labeller(spectral_var_names)) +
    theme_bw() +
    theme(legend.position = 'none',
          axis.title = element_blank(),
          axis.text.x = element_text(face = 'bold', size = 14),
          axis.text.y = element_text(size = 14),
          strip.text = element_text(face = 'bold', size = 14))
ggsave(here('figures', 'extra_figures', 'field.cluster.spectral.boxplots.png'), height = 8, width = 9)
```

#### Main
##### Wrangling
```{r}
# Tukey-Kramer Results
field.cluster <- c(rep(c(3, 1, 2), 8))
cluster.no <- c(rep(c(0.6, 1, 1), 8))
var <- c(rep('grass.pct', 3), rep('shrub.pct', 3),
               rep('seedling.pct', 3), rep('regen.density', 3),
               rep('delta_nbr', 3), rep('pct_recovery', 3),
               rep('postfire_fitted_slope', 3), rep('relative_regrowth', 3))
y.val <- c(85, 43, 32, 42, 90, 56,
           10, 14, 58, 1500, 1500, 18500, 
          -125, -20, -160, 1, 1.75, 1.6,
          33, 59, 45, 0.8, 2.5, 1)
label <- c('b', 'a', 'a', 'a', 'b', 'a', 'a', 'a', 'b', 'a', 'a', 'b',
           'a', 'b', 'ab', 'a', 'b', 'b', 'a', 'b', 'ab', 'a', 'b', 'ab')
panel.y.val <- c(rep(85, 3), rep(88, 3),
           rep(66, 3), rep(18500, 3), 
           rep(9, 3), rep(1.8, 3),
           rep(60, 3), rep(4.4, 3))
panel.label <- c('i', '', '', 'h', '', '', 'g', '', '', 'f', '', '', 'b', '', '', 'c', '', '', 'd', '', '', 'e', '', '')
tk.field.cluster.df <- data.frame(field.cluster, cluster.no, var, y.val, label, panel.y.val, panel.label) %>%     
  mutate(field.cluster = case_when(field.cluster == 1 ~ 'Shrub\ndominated',
                                   field.cluster == 2 ~ 'Tree\ndominated',
                                   field.cluster == 3 ~ 'Grass\ndominated')) 

df <- field.cluster.df %>% 
  mutate(field.cluster = case_when(field.cluster == 1 ~ 'Shrub\ndominated',
                                   field.cluster == 2 ~ 'Tree\ndominated',
                                   field.cluster == 3 ~ 'Grass\ndominated')) %>% 
  pivot_longer(cols = c(grass.pct, shrub.pct, seedling.pct, regen.density, delta_nbr, pct_recovery, postfire_fitted_slope, relative_regrowth), values_to = 'val', names_to = 'var')
```

##### Linear Y-Axis
```{r}
vars <- c('delta_nbr', 'pct_recovery', 'postfire_fitted_slope',
                       'relative_regrowth', 'seedling.pct', 
                       'shrub.pct', 'grass.pct')

plot_list <- list()

for(i in 1:length(vars)){
text.df <- tk.field.cluster.df %>% 
  filter(var == vars[i])

plot <- df %>% 
  filter(var == vars[[i]]) %>% 
  ggplot(aes(x = field.cluster, y = val, color = as.factor(field.cluster))) +
    geom_boxplot(outlier.shape = NA, alpha = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.3) +
    geom_text(aes(x = field.cluster, y = y.val, label = label),
              color = 'black', size = 8.5, fontface = 'bold',
              data = text.df)  +
    geom_text(aes(x = cluster.no, y = panel.y.val, label = panel.label),
              color = 'black', size = 18, fontface = 'bold',
              data = text.df)  +
    scale_color_manual(values = c('#DDC49B', '#002F00', '#08A045')) +
    labs(x = 'Field data cluster', y = ' ') +
    facet_wrap(~var, scales = 'free', labeller = as_labeller(spectral_field_var_names), strip.position = 'left') +
    theme_bw() +
    theme(text = element_text(family = 'Arial'),
          legend.position = 'none',
          axis.title = element_blank(),
          axis.text.x = element_text(face = 'bold', size = 17),
          axis.text.y = element_text(size = 18),
          strip.text = element_text(face = 'bold', size = 24),
          strip.background = element_blank(),
          strip.placement = 'outside',
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
plot_list[[i]] <- plot
}
```

##### Log-Transformed Y-Axis
```{r}
text.df <- tk.field.cluster.df %>% 
  filter(var == 'regen.density')

plot <- df %>% 
  filter(var == 'regen.density') %>% 
  ggplot(aes(x = field.cluster, y = val, color = as.factor(field.cluster))) +
    geom_boxplot(outlier.shape = NA, alpha = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.3) +
    geom_text(aes(x = field.cluster, y = y.val, label = label),
              color = 'black', size = 8.5, fontface = 'bold',
              data = text.df)  +
    geom_text(aes(x = cluster.no, y = panel.y.val, label = panel.label),
              color = 'black', size = 18, fontface = 'bold',
              data = text.df)  +
    scale_color_manual(values = c('#DDC49B', '#002F00', '#08A045')) +
    labs(x = 'Field Data Cluster', y = ' ') +
    facet_wrap(~var, scales = 'free', labeller = as_labeller(spectral_field_var_names), strip.position = 'left') +
    scale_y_continuous(trans = 'log', breaks = c(50, 1000, 20000)) +
    theme_bw() +
    theme(text = element_text(family = 'Arial'),
          legend.position = 'none',
          axis.title = element_blank(),
          axis.text.x = element_text(face = 'bold', size = 17),
          axis.text.y = element_text(size = 18),
          strip.text = element_text(face = 'bold', size = 24, hjust = 1),
          strip.background = element_blank(),
          strip.placement = 'outside',
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
plot_list[[8]] <- plot

#rm(plot, df, text.df)
```

##### Arranged
```{r, warning = FALSE}
field.cluster.boxplots <- cowplot::plot_grid(plot_list[[1]], plot_list[[2]], plot_list[[3]],
                                        plot_list[[4]], plot_list[[8]], plot_list[[5]],
                                        plot_list[[6]], plot_list[[7]], ncol = 4)
field.cluster.boxplots

#rm(plot_list)
```

### Bottom-Up Figure (Arranged)
```{r, warning = FALSE}
cowplot::plot_grid(field.cluster.time.series, field.cluster.boxplots, rel_heights = c(11, 9), ncol = 1)
ggsave(here('figures', 'main_figures', 'bottom.up.figure.png'), width = 22, height = 20)
```

## Stats Tests
### Clusters Diverging
```{r}
for(i in 9:23) {
  df <- long.field.cluster.df %>% 
    filter(relative_year == i)
  mod <- lmer(yearly_delta_nbr ~ field.cluster + (1|fire), df)
  print(i)
  print(emmeans(mod, pairwise ~ field.cluster, adjust = 'tukey'))
}
```

### Diffferences Between Clusters
```{r}
mod <- lmer(grass.pct ~ as.factor(field.cluster) + (1|fire), field.cluster.df)
emmeans(mod, list(pairwise ~ field.cluster), adjust = "tukey")

mod <- lmer(shrub.pct ~ as.factor(field.cluster) + (1|fire), field.cluster.df)
emmeans(mod, list(pairwise ~ field.cluster), adjust = "tukey")

mod <- lmer(regen.density ~ as.factor(field.cluster) + (1|fire), field.cluster.df)
emmeans(mod, list(pairwise ~ field.cluster), adjust = "tukey")

mod <- lmer(seedling.pct ~ as.factor(field.cluster) + (1|fire), field.cluster.df)
emmeans(mod, list(pairwise ~ field.cluster), adjust = "tukey")

mod <- lmer(pct_recovery ~ as.factor(field.cluster) + (1|fire), field.cluster.df)
emmeans(mod, list(pairwise ~ field.cluster), adjust = "tukey")

mod <- lmer(relative_regrowth ~ as.factor(field.cluster) + (1|fire), field.cluster.df)
emmeans(mod, list(pairwise ~ field.cluster), adjust = "tukey")

mod <- lmer(delta_nbr ~ as.factor(field.cluster) + (1|fire), field.cluster.df)
emmeans(mod, list(pairwise ~ field.cluster), adjust = "tukey")

mod <- lmer(postfire_fitted_slope ~ as.factor(field.cluster) + (1|fire), field.cluster.df)
emmeans(mod, list(pairwise ~ field.cluster), adjust = "tukey")
```


# --------------------

# Ridgeline Plots
```{r}
field.rect <- data.frame (xmin=-Inf, xmax=Inf, ymin=2.75, ymax=8.75)
traj.rect <- data.frame (xmin=-Inf, xmax=Inf, ymin=8.75, ymax=14)
topo.rect <- data.frame (xmin=-Inf, xmax=Inf, ymin=0, ymax=2.75)

# Using Spectral Trajectory Clusters
ridgeline.df <- field.cluster.df %>% 
  mutate(`% Grass` = scale(grass.pct),
         `% Forb` = scale(herb.pct),
         `% Shrub` = scale(shrub.pct),
         `% Juvenile conifer` = scale(seedling.pct),
         `Log regen. density` = scale(log.regen.density),
         `Canopy closure` = scale(densiometer),
         `Relative regrowth` = scale(relative_regrowth),
         `% Recovery` = scale(pct_recovery),
         `Change in NBR` = scale(delta_nbr),
         `Post-fire fitted slope` = scale(postfire_fitted_slope),
         `Burn severity (dNBR)` = scale(dnbr),
         `Elevation` = scale(elevation),
         `Slope` = scale(slope)) %>% 
  pivot_longer(cols = c(`% Grass`, `% Forb`, `% Shrub`, `% Juvenile conifer`, `Log regen. density`, `Canopy closure`, `Relative regrowth`,
                        `% Recovery`, `Change in NBR`, `Post-fire fitted slope`, `Burn severity (dNBR)`,  `Elevation`, `Slope`), names_to = 'var', values_to = 'val') %>%
  mutate(cluster = case_when(
    cluster == 2 ~ 'Fast recovery',
    cluster == 1 ~ 'Slow recovery'
  )) %>% 
  mutate(var=factor(var)) %>% 
  mutate(var=fct_relevel(var,c("Elevation", "Slope", "% Grass", "% Forb", "% Shrub", "% Juvenile conifer", "Log regen. density", "Canopy closure", "Relative regrowth", "% Recovery", "Change in NBR", "Post-fire fitted slope", "Burn severity (dNBR)"))) %>%
  filter(val < 5) %>% 
  arrange(var) 

# Significant differences
ridgeline.sig.df <- data.frame(var = c("Elevation", "Slope", "% Grass", "% Forb", "% Shrub", "% Juvenile conifer", "Log regen. density", "Canopy closure", "Relative regrowth", "% Recovery", "Change in NBR", "Post-fire fitted slope", "Burn severity (dNBR)"), sig = c('', '', '*', '', '***', '', '', '', '***', '***', '***', '***', '**'), x = c(rep(-3, 13)))

ridgeline.df %>% 
  ggplot() +
    geom_density_ridges(aes(x = val, y = var, fill = cluster), alpha = 0.75, size = 1, scale = 0.8, rel_min_height = 0.001) +
    geom_rect(data=field.rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="#00E257", alpha=0.1, inherit.aes = FALSE) +
    geom_rect(data=traj.rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="#50AAFF", alpha=0.1, inherit.aes = FALSE) +
    geom_rect(data=topo.rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="#B48B7D", alpha=0.1, inherit.aes = FALSE) +
    geom_text(data = ridgeline.sig.df, aes(label = sig, x = x, y = var), size = 10) +
    scale_fill_manual(values = c('black', 'gray70')) +
    theme_bw() +
    labs(fill = 'Spectral trajectory cluster') +
    theme(text = element_text(family = 'Arial'),
          legend.position = 'top',
          axis.title = element_blank(),
          axis.text.y = element_text(face = 'bold', size = 26),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 24),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggsave(here('figures', 'supp_figures', 'spectral.cluster.ridgeline.png'), height = 11.5, width = 14)
```

## Abridged No. 1 (For Top-Down Figure)
```{r}
field.rect <- data.frame (xmin=-Inf, xmax=Inf, ymin=2.75, ymax=8.75)
traj.rect <- data.frame (xmin=-Inf, xmax=Inf, ymin=8.75, ymax=10)
topo.rect <- data.frame (xmin=-Inf, xmax=Inf, ymin=0, ymax=2.75)
ridgeline.sig.df <- data.frame(var = c("Elevation", "Slope", "% Grass", "% Forb", "% Shrub", "% Juvenile conifer", "Log-Regen Density", "Canopy Closure", "Burn Severity (dNBR)"), sig = c('', '', '', '*', '*', '', '*', '', ''), x = c(rep(-3, 9)))

ridgeline_plot <- ridgeline.df %>% 
  filter(var %in% c("Elevation", "Slope", "% Grass", "% Forb", "% Shrub", "% Juvenile conifer", "Log-Regen Density", "Canopy Closure", "Burn Severity (dNBR)")) %>% 
  ggplot() +
    geom_density_ridges(aes(x = val, y = var, fill = cluster), alpha = 0.75, size = 1, scale = 0.8, rel_min_height = 0.001) +
    geom_rect(data=field.rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="#00E257", alpha=0.1, inherit.aes = FALSE) +
    geom_rect(data=traj.rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="#50AAFF", alpha=0.1, inherit.aes = FALSE) +
    geom_rect(data=topo.rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="#B48B7D", alpha=0.1, inherit.aes = FALSE) +
    geom_text(data = ridgeline.sig.df, aes(label = sig, x = x, y = var), size = 10) +
    scale_fill_manual(values = c('black', 'gray70')) +
    theme_bw() +
    labs(fill = 'Spectral Trajectory Cluster', x = 'Scaled Value') +
    theme(legend.position = 'top',
          axis.title.y = element_blank(),
          axis.title.x = element_text(face = 'bold', size = 18),
          axis.text = element_text(face = 'bold', size = 20),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ridgeline_plot
```

## Abridged No. 2 (For Presentation)
```{r}
field.rect <- data.frame (xmin=-Inf, xmax=Inf, ymin=0, ymax=4.75)
traj.rect <- data.frame (xmin=-Inf, xmax=Inf, ymin=4.75, ymax=7)

ridgeline.df %>% 
  filter(var %in% c("% Grass", "% Shrub", "% Juvenile conifer", "Log-Regen Density", "% Recovery", "Postfire Fitted Slope")) %>% 
  ggplot(aes(x = val, y = var, fill = cluster)) +
    geom_density_ridges(alpha = 0.75, size = 1, scale = 0.8, rel_min_height = 0.001) +
    geom_rect(data=field.rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="#00E257", alpha=0.1, inherit.aes = FALSE) +
    geom_rect(data=traj.rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="#50AAFF", alpha=0.1, inherit.aes = FALSE) +
    scale_fill_manual(values = c('black', 'gray70')) +
    theme_bw() +
    labs(fill = 'Spectral Trajectory Cluster') +
    theme(legend.position = 'top',
          axis.title = element_blank(),
          axis.text = element_text(face = 'bold', size = 34),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 30),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
ggsave(here('figures', 'extra_figures', 'abridged.spectral.cluster.ridgeline.png'), height = 8, width = 14)
```

## K-S Tests
To test whether the two clusters share similar *distributions*, we use the K-S tests; if 'sig.' then they come from different distributions, but if 'n.s.', then they come from statistically similar distributions

Data Wrangling: getting two datsets, one for each cluster
```{r}
clust1.df <- field.cluster.df %>% 
  filter(cluster == 1)
clust2.df <- field.cluster.df %>% 
  filter(cluster == 2)
```

Field Data
```{r}
ks.test(clust1.df$shrub.pct, clust2.df$shrub.pct) # sig.
ks.test(clust1.df$seedling.pct, clust2.df$seedling.pct) # n.s.
ks.test(clust1.df$grass.pct, clust2.df$grass.pct) # sig.
ks.test(clust1.df$herb.pct, clust2.df$herb.pct) # n.s.
ks.test(clust1.df$green.pct, clust2.df$green.pct) # sig.
ks.test(clust1.df$npv.pct, clust2.df$npv.pct) # n.s.
ks.test(clust1.df$log.seedling.density, clust2.df$log.seedling.density) # n.s.
ks.test(clust1.df$log.regen.density, clust2.df$log.regen.density) # n.s.
ks.test(clust1.df$densiometer, clust2.df$densiometer) # n.s.
```

Spectral Metrics and dNBR
```{r}
ks.test(clust1.df$delta_nbr, clust2.df$delta_nbr) # sig.
ks.test(clust1.df$postfire_fitted_slope, clust2.df$postfire_fitted_slope
        ) # sig.
ks.test(clust1.df$pct_recovery, clust2.df$pct_recovery) # sig.
ks.test(clust1.df$relative_regrowth, clust2.df$relative_regrowth) # sig.
ks.test(clust1.df$dnbr, clust2.df$dnbr) # n.s.
```

Climate and Topography
```{r}
ks.test(clust1.df$slope, clust2.df$slope) # n.s.
ks.test(clust1.df$elevation, clust2.df$elevation
        ) # n.s.
ks.test(clust1.df$aspect, clust2.df$aspect) # n.s.
ks.test(clust1.df$hli, clust2.df$hli) # n.s.
ks.test(clust1.df$postfire_ppt, clust2.df$postfire_ppt) # n.s.
ks.test(clust1.df$postfire_tmax, clust2.df$postfire_tmax) # n.s.
ks.test(clust1.df$postfire_tmin, clust2.df$postfire_tmin) # n.s.
ks.test(clust1.df$postfire_tmean, clust2.df$postfire_tmean) # n.s.
ks.test(clust1.df$postfire_vpdmax, clust2.df$postfire_vpdmax) # n.s.
ks.test(clust1.df$cwd_present, clust2.df$cwd_present) # sig.
```

# Writing CSV Files
```{r}
write_csv(field.cluster.df, file = here('data', 'processed_data', 'field.cluster.df.csv'))
write_csv(long.field.cluster.df, file = here('data', 'processed_data', 'long.field.cluster.df.csv'))
write_csv(wide.cluster.df, file = here('data', 'processed_data', 'wide.cluster.df.csv'))
```