---
title: 'NWCASC: Stratified Random Sampling of Field Sites'
author: "Joe Celebrezze"
date: "2023-05-03"
output: html_document
---

!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
THIS CODE WILL NOT RUN WITHOUT THE PROPER DATA DOWNLOADED ON YOUR LOCAL DRIVE. I COULD NOT PUSH ANYTHING TO GITHUB >100MB, SO CLIMATE DATA, LARGE NBR DATA (E.G., EGLEY), LANDFIRE DATA (forest cover) ARE NOT ON THE GITHUB REPOSITORY.
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
here = here::here
source(here("scripts", "source_code.R"))
library(ggpp)
library(ggfortify)
library(ggpmisc)
library(leaflet)
library(corrplot)
library(factoextra)
library(cluster)
library(ggpubr)
library(sp)
library(terra)
extract = raster::extract
```

Note: due to a lack of accessibility, Silvies River was removed from consideration below

## Rasters
```{r}
blue.mtns.ras <- raster(here('data',  'blue_mtns_misc', 'bluemtnsras.tif'))
# Postfire Regen. Data (Davis et al)
postfire.regen <- raster(here('data',  'regen_prob', 'postfire_regen_prob.tif'))
# Canopy Cover Data
canopy.cover <- raster(here('data',  'site_selection', 'LANDFIRE_2001_Canopy_Cover', 'us_105cc.tif'))
# Vegetation Height Data
veg.height <- raster(here('data',  'site_selection', 'LANDFIRE_2001_Veg_Height', 'us_105evh.tif'))
# MTBS Severity Data
egley.fire <- raster(here('data',  'site_selection', 'egley_severity.tif'))
HR.fire <- raster(here('data',  'site_selection', 'hash_rock_severity.tif'))
easy.fire <- raster(here('data',  'site_selection', 'easy_severity.tif'))
calamity.fire <- raster(here('data',  'site_selection', 'calamity_severity.tif'))
flagtail.fire <- raster(here('data',  'site_selection', 'flagtail_severity.tif'))
wheeler.fire <- raster(here('data',  'site_selection', 'wheeler_severity.tif'))
all.fires <- raster(here('data',  'site_selection', 'all_fires_severity.tif'))
all.fires.SpatRas <- terra::rast(here('data',  'site_selection', 'all_fires_severity.tif'))
# NBR Data
nbr_egley <- raster::stack(here('data',  'site_selection', 'Egley_nbr.tif'))
nbr_HR <- raster::stack(here('data',  'site_selection', 'HR_nbr.tif'))
nbr_easy <- raster::stack(here('data',  'site_selection', 'Easy_nbr.tif'))
nbr_calamity <- raster::stack(here('data',  'site_selection', 'Calamity_nbr.tif'))
nbr_flagtail <- raster::stack(here('data',  'site_selection', 'Flagtail_nbr.tif'))
nbr_wheeler <- raster::stack(here('data',  'site_selection', 'Wheeler_nbr.tif'))
# CHILI Data
chili_egley <- raster(here('data',  'site_selection', 'CHILI_Egley.tif'))
chili_HR <- raster(here('data',  'site_selection', 'CHILI_HashRock.tif'))
chili_easy <- raster(here('data',  'site_selection', 'CHILI_Easy.tif'))
chili_calamity <- raster(here('data',  'site_selection', 'CHILI_Calamity.tif'))
chili_flagtail <- raster(here('data',  'site_selection', 'CHILI_Flagtail.tif'))
chili_wheeler <- raster(here('data',  'site_selection', 'CHILI_WheelerPoint.tif'))
```

# Data Wrangling
## Projecting Rasters
```{r}
# CHILI Data
chili_egley <- projectRaster(from = chili_egley, to = egley.fire, method = 'ngb')
chili_HR <- projectRaster(from = chili_HR, to = HR.fire, method = 'ngb')
chili_easy <- projectRaster(from = chili_easy, to = easy.fire, method = 'ngb')
chili_calamity <- projectRaster(from = chili_calamity, to = calamity.fire, method = 'ngb')
chili_flagtail <- projectRaster(from = chili_flagtail, to = flagtail.fire, method = 'ngb')
chili_wheeler <- projectRaster(from = chili_wheeler, to = wheeler.fire, method = 'ngb')

# Canopy Cover
canopy.cover.egley <- projectRaster(from = canopy.cover, to = egley.fire, method = 'ngb')
canopy.cover.HR <- projectRaster(from = canopy.cover, to = HR.fire, method = 'ngb')
canopy.cover.easy <- projectRaster(from = canopy.cover, to = easy.fire, method = 'ngb')
canopy.cover.calamity <- projectRaster(from = canopy.cover, to = calamity.fire, method = 'ngb')
canopy.cover.flagtail <- projectRaster(from = canopy.cover, to = flagtail.fire, method = 'ngb')
canopy.cover.wheeler <- projectRaster(from = canopy.cover, to = wheeler.fire, method = 'ngb')

# Veg Height
veg.height.egley <- projectRaster(from = veg.height, to = egley.fire, method = 'ngb')
veg.height.HR <- projectRaster(from = veg.height, to = HR.fire, method = 'ngb')
veg.height.easy <- projectRaster(from = veg.height, to = easy.fire, method = 'ngb')
veg.height.calamity <- projectRaster(from = veg.height, to = calamity.fire, method = 'ngb')
veg.height.flagtail <- projectRaster(from = veg.height, to = flagtail.fire, method = 'ngb')
veg.height.wheeler <- projectRaster(from = veg.height, to = wheeler.fire, method = 'ngb')
```

## Random Selection of High Severity Burn Points
For fires w/ >1000 pixels in high severity burn, randomly sample 1000 to decrease computation time; for fires with <1000 pixels, use all that they have
```{r}
# so it's reproducible
set.seed(16)
crs(postfire.regen)
egley.fire <- projectRaster(egley.fire, crs = crs(postfire.regen))
egley.fire.df <- as.data.frame(rasterToPoints(egley.fire)) %>% 
  filter(egley_severity == 4) # only high severity
egley.fire.df <- sample_n(egley.fire.df, 1000) # sample 1000 points
xy <- egley.fire.df[,c(1,2)] # define coordinates
egley_pts <- egley.fire.df %>% 
  select(-egley_severity) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

easy.fire <- projectRaster(easy.fire, crs = crs(postfire.regen))
easy.fire.df <- as.data.frame(rasterToPoints(easy.fire)) %>% 
  filter(easy_severity == 4) # only high severity
easy.fire.df <- sample_n(easy.fire.df, 1000) # sample 1000 points
xy <- easy.fire.df[,c(1,2)] # define coordinates
easy_pts <- easy.fire.df %>% 
  select(-easy_severity) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

calamity.fire <- projectRaster(calamity.fire, crs = crs(postfire.regen))
calamity.fire.df <- as.data.frame(rasterToPoints(calamity.fire)) %>% 
  filter(calamity_severity == 4) # only high severity
xy <- calamity.fire.df[,c(1,2)] # define coordinates
calamity_pts <- calamity.fire.df %>% 
  select(-calamity_severity) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points
  
flagtail.fire <- projectRaster(flagtail.fire, crs = crs(postfire.regen))
flagtail.fire.df <- as.data.frame(rasterToPoints(flagtail.fire)) %>% 
  filter(flagtail_severity == 4) # only high severity
flagtail.fire.df <- sample_n(flagtail.fire.df, 1000) # sample 1000 points
xy <- flagtail.fire.df[,c(1,2)] # define coordinates
flagtail_pts <- flagtail.fire.df %>% 
  select(-flagtail_severity) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

HR.fire <- projectRaster(HR.fire, crs = crs(postfire.regen))
HR.fire.df <- as.data.frame(rasterToPoints(HR.fire)) %>% 
  filter(hash_rock_severity == 4) # only high severity
xy <- HR.fire.df[,c(1,2)] # define coordinates
HR_pts <- HR.fire.df %>% 
  select(-hash_rock_severity) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

wheeler.fire <- projectRaster(wheeler.fire, crs = crs(postfire.regen))
wheeler.fire.df <- as.data.frame(rasterToPoints(wheeler.fire)) %>% 
  filter(wheeler_severity == 4) # only high severity
wheeler.fire.df <- sample_n(wheeler.fire.df, 1000) # sample 1000 points
xy <- wheeler.fire.df[,c(1,2)] # define coordinates
wheeler_pts <- wheeler.fire.df %>% 
  select(-wheeler_severity) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points
```


## NBR Data
```{r, warning = FALSE}
raster::plot(egley.fire)
plot(egley_pts, add=TRUE)

#change band names
names(nbr_egley) <- as.character(c(1984:2021))
names(nbr_easy) <- as.character(c(1984:2021))
names(nbr_calamity) <- as.character(c(1984:2021))
names(nbr_flagtail) <- as.character(c(1984:2021))
names(nbr_HR) <- as.character(c(1984:2021))
names(nbr_wheeler) <- as.character(c(1984:2021))

# extract values from the rasters using the generated points
vals_egley <- as.data.frame(extract(nbr_egley, egley_pts)) %>% 
  mutate(fire = 'egley')
vals_easy <- as.data.frame(extract(nbr_easy, easy_pts)) %>% 
  mutate(fire = 'easy')
vals_calamity <- as.data.frame(extract(nbr_calamity, calamity_pts)) %>% 
  mutate(fire = 'calamity')
vals_flagtail <- as.data.frame(extract(nbr_flagtail, flagtail_pts)) %>% 
  mutate(fire = 'flagtail')
vals_wheeler <- as.data.frame(extract(nbr_wheeler, wheeler_pts)) %>% 
  mutate(fire = 'wheeler')
vals_HR <- as.data.frame(extract(nbr_HR, HR_pts)) %>% 
  mutate(fire = 'HR')

# add x and y coordinates to nbr values dataframes
coords_egley <- as.data.frame(egley_pts)
vals_egley <- cbind(coords_egley, vals_egley)
coords_easy <- as.data.frame(easy_pts)
vals_easy <- cbind(coords_easy, vals_easy)
coords_calamity <- as.data.frame(calamity_pts)
vals_calamity <- cbind(coords_calamity, vals_calamity)
coords_flagtail <- as.data.frame(flagtail_pts)
vals_flagtail <- cbind(coords_flagtail, vals_flagtail)
coords_wheeler <- as.data.frame(wheeler_pts)
vals_wheeler <- cbind(coords_wheeler, vals_wheeler)
coords_HR <- as.data.frame(HR_pts)
vals_HR <- cbind(coords_HR, vals_HR) %>% 
  na.omit() # For some reason, there is a bunch of NAs in the HR nbr raster stack
```

## Removing non-forest pixels
Threshold of >=25% canopy cover used, >= 5m average height

Note: veg height is in classes and the classes represent the following
109 = forest 5-10m
110 = forest 10-25m
111 = forest 25-50m
112 = forest >50m
```{r}
cc_egley <- as.data.frame(extract(canopy.cover.egley, egley_pts))
evh_egley <- as.data.frame(extract(veg.height.egley, egley_pts))
vals_egley <- cbind(vals_egley, cc_egley, evh_egley) %>% 
  rename(canopy_cover = `extract(canopy.cover.egley, egley_pts)`) %>% 
  rename(veg_height = `extract(veg.height.egley, egley_pts)`) %>% 
  filter(canopy_cover >= 25) %>% 
  filter(veg_height %in% c(109, 110, 111, 112)) # Note (reiterated from above): veg height is in classes and the classes represent the following - 109 = forest 5-10m, 110 = forest 10-25m, 111 = forest 25-50m, 112 = forest >50m

cc_easy <- as.data.frame(extract(canopy.cover.easy, easy_pts))
evh_easy <- as.data.frame(extract(veg.height.easy, easy_pts))
vals_easy <- cbind(vals_easy, cc_easy, evh_easy) %>% 
  rename(canopy_cover = `extract(canopy.cover.easy, easy_pts)`) %>% 
  rename(veg_height = `extract(veg.height.easy, easy_pts)`) %>% 
  filter(canopy_cover >= 25) %>% 
  filter(veg_height %in% c(109, 110, 111, 112))

cc_calamity <- as.data.frame(extract(canopy.cover.calamity, calamity_pts))
evh_calamity <- as.data.frame(extract(veg.height.calamity, calamity_pts))
vals_calamity <- cbind(vals_calamity, cc_calamity, evh_calamity) %>% 
  rename(canopy_cover = `extract(canopy.cover.calamity, calamity_pts)`) %>% 
  rename(veg_height = `extract(veg.height.calamity, calamity_pts)`) %>% 
  filter(canopy_cover >= 25) %>% 
  filter(veg_height %in% c(109, 110, 111, 112))

cc_flagtail <- as.data.frame(extract(canopy.cover.flagtail, flagtail_pts))
evh_flagtail <- as.data.frame(extract(veg.height.flagtail, flagtail_pts))
vals_flagtail <- cbind(vals_flagtail, cc_flagtail, evh_flagtail) %>% 
  rename(canopy_cover = `extract(canopy.cover.flagtail, flagtail_pts)`) %>% 
  rename(veg_height = `extract(veg.height.flagtail, flagtail_pts)`) %>% 
  filter(canopy_cover >= 25) %>% 
  filter(veg_height %in% c(109, 110, 111, 112))

cc_wheeler <- as.data.frame(extract(canopy.cover.wheeler, wheeler_pts))
evh_wheeler <- as.data.frame(extract(veg.height.wheeler, wheeler_pts))
vals_wheeler <- cbind(vals_wheeler, cc_wheeler, evh_wheeler) %>% 
  rename(canopy_cover = `extract(canopy.cover.wheeler, wheeler_pts)`) %>% 
  rename(veg_height = `extract(veg.height.wheeler, wheeler_pts)`) %>% 
  filter(canopy_cover >= 25) %>% 
  filter(veg_height %in% c(109, 110, 111, 112))

xy <- vals_HR[,c(1,2)] # define coordinates
HR_pts <- vals_HR %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points
cc_HR <- as.data.frame(extract(canopy.cover.HR, HR_pts))
evh_HR <- as.data.frame(extract(veg.height.HR, HR_pts))
vals_HR <- cbind(vals_HR, cc_HR, evh_HR) %>% 
  rename(canopy_cover = `extract(canopy.cover.HR, HR_pts)`) %>% 
  rename(veg_height = `extract(veg.height.HR, HR_pts)`) %>% 
  filter(canopy_cover >= 25) %>% 
  filter(veg_height %in% c(109, 110, 111, 112))

# merge the individual fires into one dataframe
vals_all_fires <- rbind(vals_egley, vals_easy, vals_calamity, vals_flagtail, vals_HR, vals_wheeler)
```


## Combining, Lengthening, Widening Datasets
```{r}
main.df <- vals_all_fires %>% 
  unite('x_y', c(x, y), sep = '_', remove = F) %>% 
  pivot_longer(cols = !c(x_y, x, y, fire, canopy_cover, veg_height), names_to = 'year', values_to = 'nbr') %>% 
  mutate(year = str_sub(year, 2)) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(relative_year = case_when( # Making column so that year = 10 is the year of fire for every fire
    fire == 'egley' ~ year - 1997, # Egley complex fire occurred in 2007
    fire == 'HR' ~ year - 1990, # HR fire: 2000
    fire == 'easy' ~ year - 1992, # Easy fire: 2002
    fire == 'calamity' ~ year - 1997, # Calamity complex fire: 2007
    fire == 'flagtail' ~ year - 1992, # Flagtail fire: 2002
    fire == 'wheeler' ~ year - 1986 # Wheeler Fire: 1996
  ))

wide.df <- main.df %>% 
  filter(relative_year >= 0 & relative_year <= 23) %>% 
  select(-year) %>% 
  pivot_wider(names_from = relative_year, values_from = nbr)
wide.df <- wide.df %>% 
  mutate(predisturbance_nbr = (`0` + `1` + `2` + `3` + `4` + `5` + `6` + `7` + `8` + `9`)/10) %>%  # Average nbr for ten years pre-fire
  mutate(postdisturbance_nbr = (`12` + `13` + `14` + `15` + `16` + `17` + `18` + `19` + `20` + `21`)/10) %>%  # Average nbr for ten years post-fire (note, year directly after fire is not included in this measurement either, as typically, the NBR for high severity burns does not seem to change much in the first year after fire)
  mutate(disturbance_nbr = (`10`+`11`)/2) %>% 
  mutate(delta_nbr = postdisturbance_nbr - predisturbance_nbr) # Avg postdisturbance nbr minus avg predisturbance nbr

long.df <- wide.df %>% 
  pivot_longer(cols = !c(x_y, x, y, fire, canopy_cover, veg_height, predisturbance_nbr, postdisturbance_nbr, disturbance_nbr, delta_nbr), names_to = 'relative_year', values_to = 'nbr') %>% 
  mutate(relative_year = as.numeric(relative_year)) %>% 
  unite('x_y', c(x, y), sep = '_', remove = F)
```

## Absolute, Relative Regrowth (Fitted)
```{r}
# Getting fitted values for nbr, using Loess model with smoothing span of 0.5
models <- long.df %>%
        tidyr::nest(-x_y) %>%
        mutate(m = purrr::map(data, loess,
                           formula = nbr ~ relative_year, span = .5),  # Perform loess calculation on each sample
              nbr_fitted = purrr::map(m, `[[`, "fitted")) # Retrieve the fitted values from each model

# Apply fitted y's as a new column
long.df <- models %>%
        select(-m) %>%
        tidyr::unnest()

# Getting correctly calculated absolute regrowth and relative regrowth metrics calculated
wide.fitted.df <- long.df %>% 
  select(x_y, x, y, relative_year, nbr_fitted) %>% 
  pivot_wider(names_from = relative_year, values_from = nbr_fitted)
wide.fitted.df <- wide.fitted.df %>% 
  mutate(absolute_regrowth = `15` - `10`) %>%  # From White et al 2017
  mutate(relative_regrowth = absolute_regrowth/(`9`-`10`))  # From White et al 2017
wide.df$absolute_regrowth <- wide.fitted.df$absolute_regrowth
wide.df$relative_regrowth <- wide.fitted.df$relative_regrowth
```

## Time to Recovery
```{r}
wide.df$max.postfire.nbr <- wide.df %>% 
  select(`11`, `12`, `13`, `14`, `15`, `16`, `17`, `18`, `19`, `20`, `21`, `22`, `23`) %>% 
  apply(1, max, na.rm=TRUE)

wide.df <- wide.df %>% 
  mutate(pct_recovery = max.postfire.nbr/predisturbance_nbr) %>% 
  filter(pct_recovery < 30) %>%  # removing crazy outlier; check to see if this affects random sampling
  mutate(time_to_0.8recovery = NA) %>% 
  mutate(time_to_0.8recovery = case_when(`11` >= 0.8*predisturbance_nbr ~ 1,
         `12` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 2,
         `13` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 3,
         `14` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 4,
         `15` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 5,
         `16` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 6,
         `17` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 7,
         `18` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 8,
         `19` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 9,
         `20` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 10,
         `21` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 11,
         `22` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 12,
         `23` >= 0.8*predisturbance_nbr & is.na(time_to_0.8recovery) ~ 13)) %>% 
  mutate(time_to_0.6recovery = NA) %>% 
  mutate(time_to_0.6recovery = case_when(`11` >= 0.6*predisturbance_nbr ~ 1,
         `12` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 2,
         `13` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 3,
         `14` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 4,
         `15` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 5,
         `16` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 6,
         `17` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 7,
         `18` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 8,
         `19` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 9,
         `20` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 10,
         `21` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 11,
         `22` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 12,
         `23` >= 0.6*predisturbance_nbr & is.na(time_to_0.6recovery) ~ 13))
```

## Using Fitted Lines
```{r}
postfire.long.df <- long.df %>% 
  filter(relative_year > 10)
prefire.long.df <- long.df %>% 
  filter(relative_year < 11) %>% 
  mutate(postfire_fitted_slope = NA) %>% 
  mutate(postfire_fitted_rsq = NA) %>% 
  select(postfire_fitted_rsq, postfire_fitted_slope, x_y, relative_year)

fitted_postfire_model <- postfire.long.df %>%
  group_by(x_y) %>%
  do(model = lm(nbr ~ relative_year, data = .))

fitted_postfire_model <- fitted_postfire_model %>% 
  mutate(postfire_fitted_slope = NA) %>% 
  mutate(postfire_fitted_rsq = NA) # Holding columns

for(i in 1:nrow(fitted_postfire_model)){
fitted_postfire_model$postfire_fitted_slope[i] <- fitted_postfire_model[[2]][[i]]$coefficients[2]
fitted_postfire_model$postfire_fitted_rsq[i] <- summary(fitted_postfire_model[[2]][[i]])$r.squared
}

fitted_postfire_model <- fitted_postfire_model %>% 
  select(-model)

postfire.long.df <- merge(fitted_postfire_model, postfire.long.df, by = 'x_y')

# Folding in new variables into long.df 
postfire.long.df <- postfire.long.df %>% 
  select(postfire_fitted_rsq, postfire_fitted_slope, x_y, relative_year)
long.df.fitted.vals <- rbind(prefire.long.df, postfire.long.df)
long.df <- merge(long.df, long.df.fitted.vals, by = c('x_y', 'relative_year'))

# Decluttering environment
rm(long.df.fitted.vals, fitted_postfire_model)
```

## Yearly Delta NBR
```{r}
# Anomaly in NBR relative to predisturbance average
long.df <- long.df %>% 
  mutate(yearly_delta_nbr = nbr - predisturbance_nbr)

long.df.avg <- long.df %>% 
  group_by(relative_year) %>% 
  summarise(avg_nbr = mean(nbr), avg_delta_nbr = mean(yearly_delta_nbr))

# AREA UNDER CURVE
auc <- function(y){
  n <- length(y)
  0.5*(y[1]+y[n]+2*sum(y[-c(1,n)]))
}
long.df <- long.df %>% 
  group_by(x_y) %>% 
  mutate(auc_delta_nbr = auc(yearly_delta_nbr))
```

## Scaled Recovery Metric, RdNBR
Relatively common in the literature, these metrics adjust NBR to be relative to predisturbance averages and the disturbance NBR values; in selecting the sites, these metrics were not utilized, but they're expected to be at least looked into in the final analyses
```{r}
long.df <- long.df %>% 
  mutate(scaled_recovery_metric = (nbr - disturbance_nbr)/(predisturbance_nbr - disturbance_nbr)) %>% 
  mutate(RdNBR = (predisturbance_nbr - nbr)/(sqrt(abs(predisturbance_nbr/1000))))
```


# Visualizations
## Principal Component Analysis
```{r}
# so it's reproducible
set.seed(16)

# Prepping Dataset
traj.metrics.df.long <- long.df %>% 
  filter(relative_year > 10) %>% 
  select(x_y, fire, postdisturbance_nbr, delta_nbr, postfire_fitted_rsq, postfire_fitted_slope, auc_delta_nbr)

traj.metrics.df.wide <- wide.df %>% 
  pivot_longer(cols = !c(x_y, x, y, fire, canopy_cover, veg_height, predisturbance_nbr, postdisturbance_nbr, disturbance_nbr, delta_nbr, absolute_regrowth, relative_regrowth, max.postfire.nbr, pct_recovery, time_to_0.8recovery, time_to_0.6recovery), names_to = 'relative_year', values_to = 'nbr') %>% 
  filter(relative_year > 10) %>% 
  select(x_y, fire, absolute_regrowth, relative_regrowth, pct_recovery)

traj.metrics.df <- merge(traj.metrics.df.long, traj.metrics.df.wide, by = c('x_y', 'fire')) %>% 
  distinct(x_y, fire, .keep_all = T) %>% 
  mutate(postfire_fitted_rsq = ifelse(is.nan(postfire_fitted_rsq), 0, postfire_fitted_rsq)) # Changing ~2 NaN's for postfire_fitted_rsq to 0 so that PCA can adequately run

traj.pca <- traj.metrics.df %>% 
  select(-x_y, -fire) %>% 
  prcomp(scale = T)

summary(traj.pca)
traj.pca$rotation[,1:2]

# Plotting correlations
traj.cor <- traj.metrics.df %>% 
  select(-x_y, -fire) %>% 
  cor()

corrplot(traj.cor, method = 'number', diag = F, type = 'lower')

# Plotting PCA (quick and easy method, since we're not really looking to present this at all)
autoplot(traj.pca, data = traj.metrics.df,
              loadings = TRUE, loadings.colour = 'black',
              loadings.label = TRUE, loadings.label.size = 4.5) +
  theme_bw() +
  theme(axis.title = element_text(face = 'bold', size = 16))

ggsave(here('figures',  'site_selection', 'traj.metrics.pca.png'), height = 8, width = 9)
```

traj.metrics.df, the dataframe tailored for the PCA is going to be used in the k-means clustering analysis

## Time Series Plots
```{r}
long.df %>% 
  ggplot(aes(x = relative_year, y = nbr, color = fire)) +
    geom_point(alpha = 0.3) +
    geom_line(alpha = 0.1) +
    geom_smooth(alpha = 0.5) +
    geom_line(data = long.df.avg, aes(x = relative_year, y = avg_nbr), color = 'black', linewidth = 2) +
    labs(x = 'Relative Year (Fire = 10)', y = 'NBR', color = 'Fire') +
    theme_bw()
ggsave(here('figures',  'site_selection', 'nbr.over.time.fires.png'), height = 8, width = 12)

long.df %>% 
  ggplot(aes(x = relative_year, y = yearly_delta_nbr, color = fire)) +
    geom_point(alpha = 0.3) +
    geom_line(alpha = 0.1) +
    geom_smooth(alpha = 0.5) +
    geom_line(data = long.df.avg, aes(x = relative_year, y = avg_delta_nbr), color = 'black', linewidth = 2) +
    labs(x = 'Relative Year (Fire = 10)', y = 'Change in NBR (Deviation from Prefire Avg)', color = 'Fire') +
    theme_bw()
ggsave(here('figures',  'site_selection', 'delta.nbr.over.time.fires.png'), height = 8, width = 12)

# To look at a single fire:
long.df.avg <- long.df %>% 
  filter(fire == 'calamity') %>% 
  group_by(relative_year) %>% 
  summarise(avg_nbr = mean(nbr), avg_delta_nbr = mean(yearly_delta_nbr))

long.df %>% 
  filter(fire == 'calamity') %>% 
  ggplot(aes(x = relative_year, y = nbr, color = fire)) +
    geom_point(alpha = 0.3) +
    geom_line(alpha = 0.1) +
    geom_smooth(alpha = 0.5) +
    geom_line(data = long.df.avg, aes(x = relative_year, y = avg_nbr), color = 'black', linewidth = 2) +
    labs(x = 'Relative Year (Fire = 10)', y = 'NBR', color = 'Fire') +
    theme_bw()
```

## Exploratory Scatterplots
```{r}
wide.df %>% 
  ggplot(aes(x = predisturbance_nbr, y = postdisturbance_nbr)) +
    geom_point() +
    stat_poly_line(color = 'black') +
    stat_poly_eq(use_label(c("eq", "R2"))) +
    labs(x = 'Prefire Avg NBR', y = 'Postfire Avg NBR') +
    theme_bw()

wide.df %>% 
  ggplot(aes(x = disturbance_nbr, y = postdisturbance_nbr)) +
    geom_point() +
    stat_poly_line(color = 'black') +
    stat_poly_eq(use_label(c("eq", "R2"))) +
    labs(x = 'Disturbance Avg NBR (2007-2008)', y = 'Postfire Avg NBR') +
    theme_bw()
```

### Using Fitted Lines (Postfire)
```{r}
long.df %>% 
  distinct(disturbance_nbr, postfire_fitted_slope, .keep_all = TRUE) %>% 
  ggplot(aes(x = disturbance_nbr, y = postfire_fitted_slope)) +
    geom_point() +
    stat_poly_line(color = 'black') +
    stat_poly_eq(use_label(c("eq", "R2"))) +
    labs(x = 'Disturbance Avg NBR (2007-2008)', y = 'Fitted Slope (from 2008-2021)') +
    theme_bw()
```

# K-Means Cluster Analysis
## Method 1: Trajectory Metrics
For this method, I am going to mirror the method used by Madeline when looking at the field data, but instead of inputting the field data values into the analysis, the trajectory metrics will be inputted and we'll get clusters for different 'modes' of post-fire recovery

Prepping dataset
```{r}
# so it's reproducible
set.seed(16)

# Only interested in the trajectory metrics
traj.df <- traj.metrics.df %>% 
  select(-x_y, -fire, -auc_delta_nbr, -absolute_regrowth, -postfire_fitted_rsq, -postdisturbance_nbr) %>% # removing some metrics that are highly correlated and some that we determined to be unimportant 
  filter(relative_regrowth > -20) # getting rid of one strange outlier 

# Scaling, centering all metrics
traj.scaled.df <- as.data.frame(scale(traj.df))

# Renaming rows
#rownames(traj.scaled.df) = traj.metrics
```

The below chunks are from Madeline's code
```{r}
# so it's reproducible
set.seed(16)

## 3 ways to determine optimal cluster 
# 1. optimal cluster at the elbow
factoextra::fviz_nbclust(traj.scaled.df, kmeans, method = "wss") # Looks like 5 to me

# 2. optimal cluster from gap statistic: high point!
gap_stat <- clusGap(traj.scaled.df,
                    FUN = kmeans,
                    nstart = 25,
                    K.max = 10,
                    B = 50)
# plot number of clusters vs. gap statistic
fviz_gap_stat(gap_stat) # Not sure from this one... Maybe 3?

# 3. silhouette analysis
fviz_nbclust(traj.scaled.df, kmeans, method = "silhouette") # Looks like 3 to me
```

```{r}
# so it's reproducible
set.seed(16)

#perform k-means clustering with k = 3 clusters
km <- kmeans(traj.scaled.df, centers = 3) # Note I removed nstart = 33, what does this mean?; note: I also tried k = 5

#view results
km

#plot results of final k-means model
fviz_cluster(km, data = traj.scaled.df)

#find unscaled (real-value) means of each cluster
aggregate(traj.df, by=list(cluster=km$cluster), mean)

#add cluster assigment to dataframes
traj.scaled.df <- cbind(traj.scaled.df, cluster = km$cluster)
traj.df <- cbind(traj.df, cluster = km$cluster)
wide.cluster.df <- merge(traj.df, wide.df, by = c('delta_nbr', 'relative_regrowth', 'pct_recovery'))
```

### Visualizations
```{r}
# From above
fviz_cluster(km, data = traj.scaled.df)
#ggsave(here('figures', 'cluster_analysis', 'cluster.plot.method1.png'), height = 8, width = 10)

# Lengthening dataframe, adding RdNBR, scaled recovery metric, and delta NBR (deviation from prefire conditions)
long.cluster.df <- wide.cluster.df %>% 
  pivot_longer(cols = !c(x_y, x, y, fire, canopy_cover, veg_height, predisturbance_nbr, postdisturbance_nbr, disturbance_nbr, delta_nbr, absolute_regrowth, relative_regrowth, max.postfire.nbr, pct_recovery, time_to_0.8recovery, time_to_0.6recovery, cluster), names_to = 'relative_year', values_to = 'nbr') %>% 
  mutate(yearly_delta_nbr = nbr - predisturbance_nbr) %>% 
  mutate(scaled_recovery_metric = (nbr - disturbance_nbr)/(predisturbance_nbr - disturbance_nbr)) %>% 
  mutate(RdNBR = (predisturbance_nbr - nbr)/(sqrt(abs(predisturbance_nbr/1000)))) %>% 
  mutate(cluster = as.factor(cluster)) %>% 
  mutate(relative_year = as.numeric(relative_year))

# Average values
long.cluster.avg.df <- long.cluster.df %>% 
  group_by(cluster, relative_year) %>% 
  summarise(avg_nbr = mean(nbr), avg_delta_nbr = mean(yearly_delta_nbr), avg_srm = mean(scaled_recovery_metric), avg_RdNBR = mean(RdNBR), se_nbr = sd(nbr)/sqrt(n()), se_delta_nbr = sd(yearly_delta_nbr)/sqrt(n()), sd_nbr = sd(nbr), sd_delta_nbr = sd(yearly_delta_nbr))

# Time series visualizations
long.cluster.df %>% 
  ggplot() +
    geom_point(aes(x = relative_year, y = nbr, color = cluster), alpha = 0.1) +
    geom_line(aes(x = relative_year, y = nbr, color = cluster), alpha = 0.05) +
    geom_line(data = long.cluster.avg.df, aes(x = relative_year, y = avg_nbr, color = cluster), alpha = 0.8, linewidth = 2) +
    geom_errorbar(data = long.cluster.avg.df, aes(x = relative_year, ymax = avg_nbr + se_nbr, ymin = avg_nbr - se_nbr, color = cluster), linewidth = 1, position = position_dodge(width = 0.15)) +
    labs(x = 'Relative Year (Fire = 10)', y = 'NBR', color = 'Cluster') +
    theme_bw()
ggsave(here('figures',  'site_selection', 'nbr.over.time.cluster.png'), height = 8, width = 12)

long.cluster.df %>% 
  ggplot() +
    geom_point(aes(x = relative_year, y = yearly_delta_nbr, color = cluster), alpha = 0.1) +
    geom_line(aes(x = relative_year, y = yearly_delta_nbr, color = cluster), alpha = 0.05) +
    geom_line(data = long.cluster.avg.df, aes(x = relative_year, y = avg_delta_nbr, color = cluster), alpha = 0.8, linewidth = 2) +
    geom_errorbar(data = long.cluster.avg.df, aes(x = relative_year, ymax = avg_delta_nbr + sd_delta_nbr, ymin = avg_delta_nbr - sd_delta_nbr, color = cluster), linewidth = 1, position = position_dodge(width = 0.15)) +
    labs(x = 'Relative Year (Fire = 10)', y = 'Change in NBR (Deviation from Prefire Avg)', color = 'Cluster') +
    theme_bw()
ggsave(here('figures',  'site_selection', 'delta.nbr.over.time.cluster.png'), height = 8, width = 12)

# Check to see if clustered by fire
wide.cluster.df %>% 
  ggplot(aes(x = fire, y = as.numeric(cluster))) +
    geom_jitter(width = 0.3, height = 0.3) +
    labs(x = 'Fire', y = 'Cluster') +
    theme_bw()
ggsave(here('figures',  'site_selection', 'clusters.vs.fires.png'), height = 8, width = 8)
```

# Investigating Clusters
See four_fires_cluster_analysis.Rmd
## Trajectory vs. Clusters Boxplots
```{r}
# Renaming facet labels
field_data_names <- c(predisturbance_nbr = "Prefire Avg. NBR",
                    postdisturbance_nbr = "Postfire Avg. NBR",
                    delta_nbr = "Change in NBR",
                    absolute_regrowth = "Absolute Regrowth",
                    relative_regrowth = 'Relative Regrowth',
                    pct_recovery = '% Recovery')

wide.cluster.df %>% 
  select(cluster, predisturbance_nbr, postdisturbance_nbr, delta_nbr, absolute_regrowth, relative_regrowth, pct_recovery) %>%  
  pivot_longer(cols = !c(cluster), names_to = 'var', values_to = 'values') %>% 
  ggplot(aes(x = as.factor(cluster), y = values)) +
    geom_boxplot() +
    facet_wrap(~var, scales = 'free', labeller = as_labeller(field_data_names)) +
    labs(x = 'Cluster #', y = ' ') +
    theme_bw() + 
    theme(axis.text = element_text(size = 14),
          axis.title = element_text(size = 16, face = 'bold'),
          strip.text = element_text(size = 15, face = 'bold'))
ggsave(here('figures',  'site_selection', 'cluster.means.boxplot.png'), height = 8, width = 12)
```



# Stratified Random Sampling
## Egley
```{r}
# so it's reproducible
set.seed(16)

egley.sites <- wide.cluster.df %>% 
  filter(fire == 'egley') %>% 
  group_by(cluster) %>% 
  slice_sample(n = 45)

xy <- egley.sites[,c(7,8)] # define coordinates
egley_sites_pts <- egley.sites %>% 
  select(x, y) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

chili_egley_df <- as.data.frame(extract(chili_egley, egley_sites_pts))
egley.sites <- cbind(egley.sites, chili_egley_df) %>% 
  rename(hli = `extract(chili_egley, egley_sites_pts)`) %>% 
  mutate(SiteID = paste0('EG_', row_number(), '_', cluster))
```

## Easy
```{r}
# so it's reproducible
set.seed(16)

easy.sites <- wide.cluster.df %>% 
  filter(fire == 'easy') %>% 
  group_by(cluster) %>% 
  slice_sample(n = 30)

xy <- easy.sites[,c(7,8)] # define coordinates
easy_sites_pts <- easy.sites %>% 
  select(x, y) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

chili_easy_df <- as.data.frame(extract(chili_easy, easy_sites_pts))
easy.sites <- cbind(easy.sites, chili_easy_df) %>% 
  rename(hli = `extract(chili_easy, easy_sites_pts)`) %>% 
  mutate(SiteID = paste0('EA_', row_number(), '_', cluster))
```

## Flagtail
```{r}
# so it's reproducible
set.seed(16)

flagtail.sites <- wide.cluster.df %>% 
  filter(fire == 'flagtail') %>% 
  group_by(cluster) %>% 
  slice_sample(n = 30)

xy <- flagtail.sites[,c(7,8)] # define coordinates
flagtail_sites_pts <- flagtail.sites %>% 
  select(x, y) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

chili_flagtail_df <- as.data.frame(extract(chili_flagtail, flagtail_sites_pts))
flagtail.sites <- cbind(flagtail.sites, chili_flagtail_df) %>% 
  rename(hli = `extract(chili_flagtail, flagtail_sites_pts)`) %>% 
  mutate(SiteID = paste0('FL_', row_number(), '_', cluster))
```

## Calamity
```{r}
# so it's reproducible
set.seed(16)

calamity.sites <- wide.cluster.df %>% 
  filter(fire == 'calamity') %>% 
  group_by(cluster) %>% 
  slice_sample(n = 20)

xy <- calamity.sites[,c(7,8)] # define coordinates
calamity_sites_pts <- calamity.sites %>% 
  select(x, y) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

chili_calamity_df <- as.data.frame(extract(chili_calamity, calamity_sites_pts))
calamity.sites <- cbind(calamity.sites, chili_calamity_df) %>% 
  rename(hli = `extract(chili_calamity, calamity_sites_pts)`) %>% 
  mutate(SiteID = paste0('CA_', row_number(), '_', cluster))
```

## Hash Rock
```{r}
# so it's reproducible
set.seed(16)

HR.sites <- wide.cluster.df %>% 
  filter(fire == 'HR') %>% 
  group_by(cluster) %>% 
  slice_sample(n = 20)

xy <- HR.sites[,c(7,8)] # define coordinates
HR_sites_pts <- HR.sites %>% 
  select(x, y) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

chili_HR_df <- as.data.frame(extract(chili_HR, HR_sites_pts))
HR.sites <- cbind(HR.sites, chili_HR_df) %>% 
  rename(hli = `extract(chili_HR, HR_sites_pts)`) %>% 
  mutate(SiteID = paste0('HR_', row_number(), '_', cluster))
```

## Wheeler Point
```{r}
# so it's reproducible
set.seed(16)

wheeler.sites <- wide.cluster.df %>% 
  filter(fire == 'wheeler') %>% 
  group_by(cluster) %>% 
  slice_sample(n = 20)

xy <- wheeler.sites[,c(7,8)] # define coordinates
wheeler_sites_pts <- wheeler.sites %>% 
  select(x, y) %>% 
  SpatialPointsDataFrame(coords = xy, proj4string = crs(postfire.regen)) # transform data to spatial points

chili_wheeler_df <- as.data.frame(extract(chili_wheeler, wheeler_sites_pts))
wheeler.sites <- cbind(wheeler.sites, chili_wheeler_df) %>% 
  rename(hli = `extract(chili_wheeler, wheeler_sites_pts)`) %>% 
  mutate(SiteID = paste0('WH_', row_number(), '_', cluster))
```

## All Sites
```{r}
all.sites <- rbind(egley.sites, easy.sites, flagtail.sites, calamity.sites, HR.sites, wheeler.sites)
write.csv(all.sites, file = here('data',  'site_selection', 'site_selection.csv'))

#NOTE: For below .csv files, I manually selected sites based on road access, limiting spatial autocorrelation (all sites are at least 300m apart (mostly 400m apart with some exceptions)), and I placed less of a priority on the sites that show recovery in the spectral trajectories. I also relabelled the sites with 'NewSiteID' which is more intuitive (i.e., when there are only 10 sites in a cluster, it will be FIRE_SITENUMBER_CLUSTERNUMBER, where SITENUMBER is between 1-10 rather than bouncing around from 1-40 or something; the previous site ID (as used in the above site selection is still in the following .csv files with the column name 'OldSiteID'))
# Main Sites
main.sites <- read.csv(here('data',  'site_selection', 'main_sites.csv'))
# Backup Sites
backup.sites <- read.csv(here('data',  'site_selection', 'backup_sites.csv'))
```


# Map
```{r}
pal <- colorFactor(c("#FFE66D", "#7D82B8", "#1C2826"), domain = c(1,2,3))
severity.pal <- colorNumeric(c('transparent', "#fdccb8", "#fc8f6f", '#f44d37', '#c5161b', 'darkred'), values(all.fires),
  na.color = "transparent")

sites.map <- leaflet() %>% 
  setView(lat = 43.7,lng = -119.4, zoom = 10) %>% 
  addProviderTiles(providers$Esri.WorldStreetMap, group = 'Street Map') %>% #Topographic: Stamen.TopOSMRelief, Street: Esri.WorldStreetMap
  addRasterImage(postfire.regen, opacity = 0.65, group = 'Postfire Recovery Prob.') %>% 
  addRasterImage(all.fires, colors = severity.pal, opacity = 0.5, group = 'Fire Severity') %>% #note: HR still shows up.. oh well
  addRasterImage(blue.mtns.ras, opacity = 0.15, group = 'Blue Mtns', color = 'black') %>% 
  addCircleMarkers(data = main.sites, lng = main.sites$x, lat = main.sites$y, radius = 4,
                   color = ~pal(cluster), popup = ~paste(NewSiteID, '<br> HLI:', hli, '<br> CC:', canopy_cover),
                   label = ~paste(NewSiteID, 'HLI:', hli, 'CC:', canopy_cover),
                   stroke = F, fillOpacity = 1,
                   group = 'Main Sites') %>% 
  addCircleMarkers(data = backup.sites, lng = backup.sites$x, lat = backup.sites$y, radius = 4,
                   color = ~pal(cluster), popup = ~paste(NewSiteID, '<br> HLI:', hli, '<br> CC:', canopy_cover),
                   label = ~paste(NewSiteID, 'HLI:', hli, 'CC:', canopy_cover),
                   stroke = F, fillOpacity = 1,
                   group = 'Backup Sites') %>% 
  addMeasure() %>% 
  addLayersControl(
    baseGroups = c('Street Map'),
    overlayGroups = c('Postfire Recovery Prob.', 'Fire Severity', 'Blue Mtns', 'Main Sites', 'Backup Sites'),
    options = layersControlOptions(collapsed = TRUE) 
  ) %>% 
  hideGroup(c('Postfire Recovery Prob.', 'Blue Mtns', 'Backup Sites'))
sites.map
```

